---
title: "Deborah_bead_data"
author: "Biplabendu Das"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    keep_md: no
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

set.seed(420)
rm(list = ls())

```

### Housekeeping

```{r housekeeping, include=FALSE}
## Load the libraries
pacman::p_load(tidyverse, viridis, conflicted)
pacman::p_load(ggplot2, patchwork, mgcv)
# pacman::p_load(RSQLite, dbplyr)

## set conflict preference
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("union", "dplyr")
conflict_prefer("lmer", "lme4")
conflict_prefer("gam", "mgcv")

## set path to your working directory
path_to_repo = "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data"
path_to_figures = paste0(path_to_repo,"/results/figures/")

## load functions
# customized theme for publication quality figures
source(paste0(path_to_repo,"/functions/theme_publication.R"))
# # function to perform enrichment analysis
# source(paste0(path_to_repo,"/functions/enrichment_analysis.R"))
# # # function to plot z-scores (Cflo genes only)
# source(paste0(path_to_repo,"/functions/plot_zscores.R"))

bd.saveplot <- function(plot,name="myplot",width,height) {
  savingto <- paste0(path_to_figures,name,".png")
  writeLines(paste0("Plot is being saved to \n",savingto))
  png(savingto, 
    width = width, height = height, units = "cm", res = 300)
  print(plot)
  trash <- dev.off()
}

# Model validation plots
### manually (as per Zurr's book, p51-p57)
### let's make a function that takes a gam model and creates these plots
mod.validate <- function(mod) {
  which.mod <- mod
  ### Independence - 1
  p1 <- s.tidydat.mod %>% mutate(residuals = resid(which.mod), fitted_values = fitted(which.mod)) %>% 
    ggplot(aes(x=cum_prop, y=residuals)) +
    geom_abline(intercept = 0, slope = 0, col="red") +
    geom_jitter(width=0.01, shape=21, size=2) +
    theme_Publication()
    #### Note: if the points are NOT evenly scattered around the red line indicates that the model violates
    ####    the normality assumption (which is likely the case sense since cum_prop data are temporally correlated)
  ### Independence - 2
  p2 <- s.tidydat.mod %>% mutate(residuals = resid(which.mod), fitted_values = fitted(which.mod)) %>% 
    ggplot(aes(x=age_class, y=residuals)) +
    geom_boxplot() +
    geom_abline(intercept = 0, slope = 0, col="red") +
    theme_Publication()
    #### Note: most residuals for 5-9 year old colonies have positive values, suggesting that the covariate "age-class"
    ####    should be added to the model (as per Zurr's book, pg. 51)
  ### Independence - 3
  p3 <- s.tidydat.mod %>% mutate(residuals = resid(which.mod), fitted_values = fitted(which.mod)) %>% 
    ggplot(aes(x=age_obs, y=residuals)) +
    geom_abline(intercept = 0, slope = 0, col="red") +
    geom_jitter(width=0.01, shape=21, size=2) +
    theme_Publication()
    #### Note:
  ### Independence - 4
  p4 <- s.tidydat.mod %>% mutate(residuals = resid(which.mod), fitted_values = fitted(which.mod)) %>% 
    ggplot(aes(x=prop_foraging_30s, y=residuals)) +
    geom_abline(intercept = 0, slope = 0, col="red") +
    geom_jitter(width=0.01, shape=21, size=2) +
    theme_Publication()
  ### Independence - 5
  p5 <- s.tidydat.mod %>% mutate(residuals = resid(which.mod), fitted_values = fitted(which.mod)) %>% 
    ggplot(aes(x=bead_color, y=residuals)) +
    geom_boxplot() +
    geom_abline(intercept = 0, slope = 0, col="red") +
    theme_Publication()
  
  ## combine the plots and return as output
  return((p1+p2)/(p3+p4)/p5)
}
```


```{r set_parameters}
## Set color scheme for age classes
age_colors <- c("#8D72A1", "#F26C7F", "#FBAD73", "#9CD5A2", "#F29544")
```

## Overview/Goals

Email from Deborah: 

This data set comes from an experiment with harvester ants last summer. Harvester ant foragers bring in lots of seeds and parts of plants, some of which are rejected by ants inside the nest. Eventually, the refuse is carried out and put on the nest mound midden. 

> Here we presented small plastic beads coated in oil which the ants brought into the nest, and were later put back out on the midden, outside the colony, close to the nest entrance.

## Bead data 2021

```{r notes_v1, eval=F, include=F, echo=F}
# ## Part 1: INSPECT DATA
# 
# Two data sheets are available in the Excel workbook. For convenience of analysis, the excel sheets are saved individually as csv files and, if necessary, the column names for each were modified.
# 
# Sheet 1: foraging_data.csv
# - contains three columns: date, colonyID, foraging_30s
# 
# Sheet 2: beads_out.csv
# - contains XXX cols: num_out, date, colonyID, time_put_out, n_collected, n_returned_DayX (X %in% 1:6)
# 
# Brief description of the columns:
# - date = date of observation
# - colonyID = unique colony identifier
# - time_put_out = time-of-day at which beads were put out
# - n_collected = number of beads that had been brought into the nest at the end of foraging on Day 1
# - n_returned_DayX = number of beads that had been brought *out of the nest* by the end of **Day X**
# 
# Note; 
# NA means the colony is not active that day
# NM means that the colony was doing nest maintenance, when they carry out refuse including beads
# 
# Some additional info:
# 
# Beads coated with oil were put out in 6 colonies on day 1, color green, then another 6 colonies on day 2, color green, then again with first 6 on day 3 color pink, then again with 2nd 6 on day 4 color pink, back to green days 5 and 6
# 
# Numbers of beads brought out are listed separately by color. E.g., colony 1131 took out 18 green beads on day 2 of green 8-24, and 3 pink beads on day 2 of pink 8-26.
# 
# For the 3rd round, for whatever number of green beads from the first round were not out by 8-27 or 8-28, then we don't know which day those were from.

```



### Load data

```{r load_data}
# read foraging_data.csv
for.dat <- read.csv(paste0(path_to_repo,"/data/foraging_data.csv"),
                    stringsAsFactors = F, header = T) %>% as_tibble()

# read bead data
bead.dat <- read.csv(paste0(path_to_repo, "/data/beads_out_new.csv"),
                stringsAsFactors = F, header = T) %>% as_tibble()

# read colony-age data
colony.dat <- read.csv(paste0(path_to_repo, "/data/foraging_counts_08_13_22.csv"),
                       stringsAsFactors = F, header = T) %>% as_tibble() %>% 
  select(colonyID = colony_ID, everything()) %>% 
  # mutate(colonyID = as.factor(colonyID)) %>% 
  mutate(age_class = ifelse(age_2022 <= 2, "1-2_yr_old",
                                  ifelse(age_2022 > 2 & age_2022 <= 4, "3-4_yr_old",
                                         "older_than_5_yr")))
```


### Summarize the data

```{r summarize_data}

writeLines("What are the names of the colonies? \nShown as: Colony-ID")
bead.dat %>% 
  select(colonyID) %>% 
  distinct() %>% 
  left_join(colony.dat, by="colonyID") %>%
  head() %>% 
  arrange(age_2022)


# Specify the dates for green/pink beads
green.days <- c("8/23/21","8/24/21", "8/27/21","8/28/21") %>% as.Date(format = "%m/%d/%y")
pink.days <- c("8/25/21","8/26/21") %>% as.Date(format = "%m/%d/%y")
```

### Clean Data

```{r clean_1}
df <-
  bead.dat %>%
  
  # retain only useful data
  na.omit() %>% 
  
  # arrange by colony IDs
  arrange(colonyID) %>% 
  
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
  mutate(colonyID = factor(colonyID)) %>% 
  
  # add a column that specifies the color of beads used
  mutate(bead_col = ifelse(date %in% green.days[1:2], "green",
                           ifelse(date %in% green.days[3:4], "green-2",
                           ifelse(date %in% pink.days, "pink", NA)))) %>%
  
  mutate(batch = ifelse(date %in% c(green.days[c(1,3)],pink.days[1]), "batch_A",
                        ifelse(date %in% c(green.days[c(2,4)],pink.days[2]), "batch_B",NA))) %>% 
  
  
  # # SIDE NOTE: Explore the original dataset
  # mutate_at(vars(colnames(.)[names(.) %in% c(paste0("n_returned_Day",1:6))]),
  #       .funs = funs(ifelse(.=="", NA, as.character(.)))) %>% 
  # mutate_at(vars(colnames(.)[names(.) %in% c(paste0("n_returned_Day",1:6))]),
  #       .funs = as.numeric) %>% head()
  
  
  # CLEAN: Keep only data for first three days of observations after presenting beads
  select(num_out:n_returned_Day3, bead_col, batch) %>%
  
  # format columns appropriately
  mutate(n_collected = as.numeric(readr::parse_number(n_collected))) %>% 
  mutate(n_returned_Day1 = as.numeric(n_returned_Day1)) %>% 
  mutate(n_returned_Day2 = as.numeric(n_returned_Day2)) %>% 
  mutate(n_returned_Day3 = as.numeric(n_returned_Day3))
  

# query 1  
# df %>% 
#   # summarize the data by batch and bead_col
#   group_by(batch, bead_col, date) %>%
#   summarize(n_colonies = n())
  
# query 2  
# It seems the data for 25-Aug-2021, Colony-1474 is missing (or too little);  
df <- 
  df %>% 
  # remove it
  filter(!is.na(n_collected))

```


CLEAN UP:

1. There are four colonies (1 green, 3 pink) that collect <30 of the 50 beads presented. [n_collected = 8, 9, 17, 20]. It might be a good idea to leave them out before we analyze the dat, at least for Colonies that collected ≤10 beads of the 50 presented. [cleanup 1]

2. Data for Colony 1288 (green-1; 24 to 26 Aug) shows that 48 beads were presented but 50 beads were collected at the refuse pile. Removing that from our dataset as I am unsure of the source of the error. [cleanup 2]


```{r remove_sample}
dat <-
  df %>%
  # remove the time column (not necessary; almost the same time) 
  select(-4) %>% 
  filter(!(bead_col=="pink" & colonyID %in% c("969","1338"))) %>% # cleanup 1
  filter(!(bead_col=="green" & colonyID == "1288")) # cleanup 

writeLines("Which colony-days remain after cleanup?")
dat %>% 
  select(date, colonyID) %>% 
  distinct() %>% 
  mutate(colonyID = as.numeric(as.character(colonyID))) %>%
  left_join(colony.dat, by="colonyID") %>% 
  group_by(colonyID, age_2022, age_class) %>% 
  summarize(n_days = n()) %>% 
  arrange(age_2022)
```


```{r prep_data}
## Part 5: Plot cum. prop.

tidydat.2021 <-
  dat %>% 
  
  na.omit() %>%
  
  ## only keep the colonies that collected more than 40 beads
  filter(n_collected >= 40) %>% 
  
  # Cumulative prop 
  # create the cumulative sums
  mutate(cum_prop_day1 = n_returned_Day1/n_collected,
         cum_prop_day2 = (n_returned_Day1+n_returned_Day2)/n_collected,
         cum_prop_day3 = (n_returned_Day1+n_returned_Day2+n_returned_Day3)/n_collected) %>% 
  
  select(num_out, date, colonyID, n_collected, bead_col, batch:cum_prop_day3) %>% 
  
  pivot_longer(cols = starts_with("cum_prop"),
               names_to = "day_of_obs",
               values_to = "cum_prop_returned") %>% 
  
  mutate(day_of_obs = as.numeric(readr::parse_number(day_of_obs))) %>% 
  mutate(date = date+day_of_obs-1)
  
# # sumarrize the tidydat
# tidydat.2021 %>% 
#   group_by(batch, bead_col, date) %>%
#   summarize(n_colonies = n())
# 
# # what proportion of collected beads do they put out by day 3?
# tidydat.2021 %>% 
#   filter(day_of_obs=="3") %>%
# # 
# #   pull(n_collected) %>% sd()/sqrt(51)
# #   sd()/sqrt(17) # 17 = length of cum_prop_returned
# 
#   group_by(bead_col, batch) %>% 
#   summarize(n_colonies=n(),mean_prop_returned=mean(cum_prop_returned))

```


___

### Fig. 1-2

```{r abc}
# Selecting old colonies

old.summ <- 
  tidydat.2021 %>% 
  mutate(colonyID = as.character(colonyID) %>% as.numeric()) %>% 
  left_join(colony.dat, by="colonyID") %>% 
  # select(date, colonyID, bead_col, day_of_obs, cum_prop_returned, age_2022) %>% 
  
  mutate(age_obs = age_2022-1) %>% 
  
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr_old",
                            ifelse(age_obs >= 5, "older_than_5_yr", "3-4_yr_old"))) %>% 
  mutate(age_class = factor(age_class, levels = c("1-2_yr_old", "3-4_yr_old", "older_than_5_yr"))) %>% 
  
  mutate(day_of_obs = paste0("D",day_of_obs,"_bN")) %>% 
  mutate(day_of_obs = factor(day_of_obs, levels = c("D1_bN", "D2_bN", "D3_bN"))) %>% 
  
  mutate(bead_col = factor(paste0(bead_col,"-21"))) %>% 
  mutate(ID = paste0(colonyID,"_",bead_col, "_", age_obs)) %>% 
  
  ## change column names to make them compatible to the 2022 dataset
  select(colonyID, date,
         bead_color = bead_col,
         age_2022, age_obs, age_class, 
         beads_collected = n_collected,
         counted_when = day_of_obs,
         cum_prop = cum_prop_returned)


old.summ %>% 
  as_tibble() %>% 
  filter(counted_when=="D3_bN") %>% 
  group_by(colonyID,age_class) %>% 
  summarize(n_reps = n()) %>% 
  right_join(old.summ %>% select(colonyID, age_obs) %>% distinct()) %>% 
  arrange(age_obs)

bd.saveplot(
  name = "01figure_bead2021",
  width = 32, height = 10,
  old.summ %>%   
   mutate(ID = paste0(colonyID,"_",bead_color, "_", age_obs)) %>% 
  ggplot() + 
  geom_line(aes(x=factor(counted_when), y=cum_prop, group=ID),
            size=2,
            alpha=0.6) +
  facet_wrap(~ID, nrow = 2) +
  theme_Publication(base_size = 15) +
  # scale_color_manual(values=c("darkgreen","darkred")) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1),
                     labels = c("0", "", "0.4","","0.8",""),
                     limits = c(0,1)) +
  xlab("Time (day of obs)") +
  ylab("Cumulutative proportion \nof beads returned") +
  theme(axis.text.x = element_text(angle = 90))
)

bd.saveplot(
  name = "02figure_bead2021",
  width = 10, height = 10,
  old.summ %>% 
    ggplot(aes(x=counted_when, y=cum_prop, 
               fill=age_class, color=age_class)) +
    geom_jitter(size=2, alpha=0.4, width=0.2) +
    # geom_line(aes(group=interaction(colonyID,age_class)), size=0.5, alpha=0.7) +
    geom_boxplot(alpha=0.3, outlier.color = NA) +
    facet_grid(~age_class) +
    theme_Publication(base_size = 12) +
    # rotate x-axis lables
    theme(axis.text.x=element_text(angle=90,hjust=1))+
    scale_fill_manual(values = age_colors) +
    scale_color_manual(values = age_colors) +
    labs(x = "Day-time of observation",
         y= "cumulative proportion \nof beads returned",
         caption = "bN = before NOON; D1 = Day 1") +
    theme(legend.position = "none")
)

```

___


## Bead data 2022


```{r data_2022}
filename <- "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data/data/bead_data_2022/all_bead_data_2022/2022_bead_data_master_replicate_one_two_three_29Aug22.csv"
dat.2022 <- read.csv(filename,
                     header = T, stringsAsFactors = F,
                     na.strings = c(""," ","did_not_count","NA",
                                    "did_not_check",
                                    "it_rained_could_not_check")) %>% 
  as_tibble() %>% 
  
  # select the relevant columns
  select(date, colonyID, age_class, age_2022, 
         bead_color, beads_put_out, beads_collected,
         beads_returned_Day1_beforenoon,
         beads_returned_Day1_afternoon,
         beads_returned_Day2_morning,
         beads_returned_Day2_beforenoon,
         beads_returned_Day2_afternoon,
         beads_returned_Day3_beforenoon,
         beads_returned_Day4_morning,
         beads_returned_Day4_beforenoon) %>% 
  
  # calculate the cumulative proportion of beads returned
  group_by(date, bead_color) %>% 
    
  # filter out the colonies for which we still need to check the data
  filter(!(colonyID=="1560" & bead_color=="green")) %>% 
  filter(!(colonyID=="1288" & bead_color=="green")) %>%
  filter(!(colonyID=="1546" & bead_color=="green")) %>%
  
  # convert all the bead count columns as numeric
  mutate_at(vars(matches("beads_")), as.numeric) %>% 
  
  ungroup()


## Create a tidy dataset and calculate the cumulative proportion of beads returned by colonies
tidydat.2022 <-
  dat.2022 %>% 
  
  # replace all the NAs with 0 to be able to calculate proportions
  mutate_at(vars(matches("beads_")), function(x) ifelse(is.na(x), 0, x)) %>%
  
  # day 1 - before noon
  mutate(cum_prop_returned_D1_beforenoon = round(beads_returned_Day1_beforenoon/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D1_afternoon = round(cum_prop_returned_D1_beforenoon+beads_returned_Day1_afternoon/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D2_morning = round(cum_prop_returned_D1_afternoon+beads_returned_Day2_morning/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D2_beforenoon = round(cum_prop_returned_D2_morning+beads_returned_Day2_beforenoon/beads_collected,2)) %>%
  mutate(cum_prop_returned_D2_afternoon = round(cum_prop_returned_D2_beforenoon+beads_returned_Day2_afternoon/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D3_beforenoon = round(cum_prop_returned_D2_beforenoon+beads_returned_Day3_beforenoon/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D4_morning = round(cum_prop_returned_D3_beforenoon+beads_returned_Day4_morning/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D4_beforenoon = round(cum_prop_returned_D4_morning+beads_returned_Day4_beforenoon/beads_collected,2)) %>% 
  
  mutate(beads_collected_group = ifelse(beads_collected < 22, "less_than_22", 
                                        ifelse(beads_collected > 40, 
                                               "greater_than_40", "between_22_and_40"))) %>%
  arrange(age_2022, beads_collected_group) %>% 
  
  select(colonyID,
         age_2022, age_class,
         beads_collected_group,
         bead_color,
         cum_prop_returned_D1_beforenoon:cum_prop_returned_D4_beforenoon,
         everything())

  
```


```{r prep_data_for_plot}
tidydat.2022 %<>% 
  
  ## keep only colonies that collected at least 40 (of 50) beads offered
  filter(beads_collected_group == "greater_than_40") %>% 
  
  pivot_longer(cols = starts_with("cum_prop"),
               names_to = "counted_when",
               values_to = "cum_prop") %>% 
  
  ## format the ordinal columns correctly
  mutate(counted_when = ifelse(counted_when == "cum_prop_returned_D1_beforenoon", "D1_bN",
                            ifelse(counted_when == "cum_prop_returned_D1_afternoon", "D1_aN",
                                   ifelse(counted_when == "cum_prop_returned_D2_morning", "D2_M",
                                          ifelse(counted_when == "cum_prop_returned_D2_beforenoon", "D2_bN",
                                            ifelse(counted_when == "cum_prop_returned_D2_afternoon", "D2_aN",
                                            ifelse(counted_when == "cum_prop_returned_D3_beforenoon", "D3_bN",
                                            ifelse(counted_when == "cum_prop_returned_D4_morning", "D4_M",
                                                 "D4_bN")))))))) %>%
  mutate(counted_when = factor(counted_when, 
                               levels = c("D1_bN", "D1_aN", 
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  mutate(bead_color = factor(ifelse(bead_color %in% c("light_blue", "dark_blue"), "blue", bead_color),
                             levels = c("pink","green","blue")))  %>% 
  mutate(colonyID = factor(colonyID)) %>% 
  mutate(date = as.Date(date, format="%m/%d/%y")) %>% 
  
  mutate(age_obs = age_2022) %>% 
  
  # ## remove the timepoints that we do not have data for  
  # pink beads (round one)
  filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  filter(!(bead_color=="pink" & counted_when == "D2_aN")) %>%
  filter(!(bead_color=="pink" & counted_when == "D4_M")) %>%
  filter(!(bead_color=="pink" & counted_when == "D4_bN")) %>%
  # green beads (round two)
  filter(!(bead_color=="green" & counted_when == "D2_aN")) %>%
  filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  filter(!(bead_color=="green" & counted_when == "D4_M")) %>%
  # blue beads (round three)
  filter(!(bead_color=="blue" & counted_when == "D1_aN")) %>%
  filter(!(bead_color=="blue" & counted_when == "D2_bN")) %>%
  filter(!(bead_color=="blue" & counted_when == "D4_bN")) %>% 
  
  ## select columns to make them compatible with 2021 dataset
  select(colonyID, date,
         bead_color,
         age_2022, age_obs, age_class, 
         beads_collected, counted_when, cum_prop)
  
```

### Combine the 2021 and 2022 datasets

> Currently showing all the data from 2021 and 2022.

```{r combine_datasets}
## Combine the two data sets
tidydat <- rbind(old.summ, tidydat.2022) %>%
  mutate(counted_when = factor(counted_when,
                               levels = c("D1_bN", "D1_aN",
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  ### 4 age classes
  # mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
  #                           ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
  #                                  ifelse(age_obs >=5 & age_obs < 10, "5-9_yr",
  #                                                "10+_yr")))) %>% 
  # mutate(age_class = factor(age_class, 
  #                           levels = c("1-2_yr", "3-4_yr",
  #                                      "5-9_yr", "10+_yr")))
  ### 3 age classes
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
                            ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
                                   "5+_yr"))) %>% 
  mutate(age_class = factor(age_class, levels = c("1-2_yr", "3-4_yr", "5+_yr")))

# # use only 2022 data
# tidydat <- tidydat.2022
```

## Clean up step

```{r alldata_cleanup}
# 1. Only keep instances where we have three data points per colony-trial, and cum-prop<=1
tidydat <-
  tidydat %>% 
  mutate(bead_color = as.character(bead_color)) %>% 
  mutate(bead_color = ifelse(is.na(bead_color), "UN",bead_color)) %>% 
  group_by(colonyID,bead_color) %>% 
  summarize(n.timepoints = n(),
            max.cumprop = max(cum_prop)) %>% 
  filter(n.timepoints >= 3) %>% 
  filter(max.cumprop <=1) %>% 
  right_join(tidydat, by=c("colonyID","bead_color")) %>% 
  filter(!is.na(n.timepoints)) %>% 
  ungroup()

# 2. Format and order columns
tidydat$bead_color = factor(tidydat$bead_color, levels = c("pink-21","green-21","pink","green","blue"))
which.colonies <- tidydat %>% select(colonyID, age_obs, age_class) %>% distinct()
tidydat <- tidydat %>% 
  mutate(colonyID = factor(colonyID, levels = which.colonies %>% arrange(age_obs) %>% pull(1) %>% unique()))
```



## Cumulative bead returned
### Fig. 3-4

```{r plot_boxplot}
## All colonies - 1
bd.saveplot(
  name="03figure_beadall",
  width = 25, height = 20,
  ##
  tidydat %>% 
  ggplot(aes(x=counted_when, y=cum_prop, fill=age_class, color=age_class, group=age_class)) +
  # geom_jitter(size=2, alpha=0.4, width = 0.2) +
  geom_line(aes(group=interaction(colonyID,bead_color)),
            size=2, alpha=0.7) +
  facet_grid(age_class~bead_color) +
  # coord_flip() +
  scale_y_continuous(n.breaks = 3,
                     labels = c("","0.5", "1")) +
  scale_x_discrete(labels = c("D1_bN", "", "", "D2_bN", "", "D3_bN", "", "")) +
  theme_Publication(base_size = 20) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1),
        legend.position = "none")+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       caption = "\nbN = before NOON, D1 = Day 1")
)

## All colonies - 2
bd.saveplot(
  name="04figure_beadall",
  width = 33, height = 25,
  ##
  tidydat %>% 
  ggplot(aes(x=counted_when, y=cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.5, width = 0.2) +
  geom_line(aes(group=bead_color), size=2, alpha=0.7) +
  facet_wrap(colonyID~bead_color, nrow=6) +
  scale_y_continuous(breaks = c(0,0.5,1), limits = c(0,1.2), labels = c("","0.5","1.0")) +
  
  theme_bw(base_size = 15) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none")+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       caption = "\nnot normalized for foraging activity")
)

```


### Potential trash accumulated in colony

> The daily trash accumulated in the colony is most likely proportional to the number of incoming foragers on a given day.


## Normalized rate of trash removal 

> cumulative number of beads returned / possible number of incoming foragers on Day 1

I am using the total number of foragers going in and out of the colony, for 30 secs, on Day 1 as a proxy for the possible number of incoming foragers.

```{r load_2022_foraging}
# set path to your working directory (change, if necessary)
path_to_repo2 = "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/01_field_season_2022/04_field_work/02_rprojects/01_checking_data"

# set path to the behavioral data within your repository 
path_to_data2 = "/01_data/"

## Specify the name of the master file as saved on your computer
filename2 <- "arizona_2022_master_list_Aug29_1440h.csv"

column.names <- c("date",
                  "was_it_raining",
                  "observer",
                  "colony_ID",
                  "first_obs",
                  "second_obs",
                  "time",
                  "in_1",
                  "out_1",
                  "in_2",
                  "out_2",
                  "in_3",
                  "out_3",
                  "in_4",
                  "out_4",
                  "in_5",
                  "out_5",
                  "in_6",
                  "out_6",
                  "in_7",
                  "out_7",
                  "in_8",
                  "out_8"
                  )

for.dat <- 
  read.table(paste0(path_to_repo2, path_to_data2, filename2),
                  sep = ",",
                  header = T,
                  na.strings = c(""," "),
                  stringsAsFactors = F,
                  skip = 1) %>% 
  as_tibble() %>% 
  mutate(date = as.POSIXct(date, format="%m/%d/%Y", tz="MST")) %>% 
  select(1:length(column.names))

colnames(for.dat) <- column.names

tidy.for.2022 <-
  for.dat %>% 
  # remove columns containing first and second obs
  select(-first_obs, -second_obs, -was_it_raining) %>% 

  group_by(date,observer,colony_ID,time) %>% 
  mutate(tot_in = sum(in_1,in_2,in_3,in_4,in_5,in_6,in_7,in_8, na.rm = T),
         tot_out = sum(out_1,out_2,out_3,out_4,out_5,out_6,out_7,out_8, na.rm=T)) %>% 
  mutate(foraging_30s = tot_in+tot_out) %>% 
  ungroup() %>% 
  
  mutate(colony_ID = as.factor(colony_ID),
         observer = as.factor(observer)) %>% 
  
  # na.omit() %>% 
  
  pivot_longer(.,
               cols = in_1:out_8,
               names_to = "inout",
               values_to = "n_ants") %>% 
  
  group_by(inout) %>% 
  mutate(inout2 = str_split_fixed(inout, "_", 2)[1]) %>%
  ungroup() %>% 
  
  select(date:time, inout, 
         inout2,
         n_ants, tot_in, tot_out, everything())

all.dates <- tidy.for.2022 %>% pull(date) %>% na.omit() %>% unique()
date.max <- tidy.for.2022 %>% pull(date) %>% na.omit() %>% max()
```


The task is to obtain the foraging (in and out) data for each of the bead colonies, for the duration of the different experimental rounds.

> pink: 8/16/2022; green: 8/22/22; blue: 8/26/22

```{r clean_foraging_data}
which.dates <- c(seq(as.Date("8/16/22", format="%m/%d/%y"), 
                    as.Date("8/19/22", format="%m/%d/%y"), by=1),
                 seq(as.Date("8/22/22", format="%m/%d/%y"), 
                     as.Date("8/25/22", format="%m/%d/%y"), by=1),
                 seq(as.Date("8/26/22", format="%m/%d/%y"), 
                     as.Date("8/29/22", format="%m/%d/%y"), by=1))
bead.dates <- data.frame(date = which.dates,
                         obs_day = factor(rep(c("D1", "D2", "D3", "D4"), 3),
                                          levels = c("D1", "D2", "D3", "D4")),
                         bead_color = factor(rep(c("pink", "green", "blue"), each=4),
                                             levels = c("pink", "green", "blue"))
                                             )
bead.colonies <- tidydat %>% pull(colonyID) %>% unique() %>% as.character()
which.day.obs <- bead.dates %>% mutate(date_obs = paste0(date, obs_day)) %>% pull(date_obs)

tidy.for <-
  tidy.for.2022 %>% 
  mutate(date = as.character(date)) %>% 
  filter(date %in% unique(as.character(bead.dates$date))) %>% 
  mutate(date = as.Date(date)) %>%
  filter(colony_ID %in% bead.colonies) %>% 
  left_join(bead.dates, by=c("date")) %>% 
  select(date, bead_color, colony_ID, obs_day, time, tot_in, tot_out, foraging_30s) %>% 
  distinct() %>% 
  arrange(bead_color, colony_ID, date) %>% 
  ungroup()

tidy.for <-
  tidy.for %>%
  # calculate maximum foraging for each colony 
  group_by(colony_ID) %>% 
    summarise(max_foraging_30s = max(foraging_30s)) %>% 
    ungroup() %>% 
    right_join(tidy.for) %>% 
    mutate(prop_foraging_30s = foraging_30s/max_foraging_30s) %>% 
    mutate_if(is.numeric, round, 2)


s.tidydat.2022 <-
  tidydat.2022 %>% 
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% 
  full_join(tidy.for %>% select(-time) %>% 
              select(colonyID=colony_ID, everything()) %>% 
              filter(obs_day == "D1"), 
            by = c("date","bead_color","colonyID")) %>% 
  na.omit() %>% 
  mutate(s_cum_prop = round(cum_prop/foraging_30s, 2)) %>% 
  mutate(s2_cum_prop = round(cum_prop/prop_foraging_30s, 2)) %>% 
  na.omit() %>% 
  # filter(s_cum_prop <= 1) %>% 
  ## make sure the date column represents the actual dates during which the bead experiment was conducted
  mutate(date = date+parse_number(as.character(counted_when))-1) %>% 
  select(colonyID:bead_color,
         age_2022:age_class, 
         beads_collected,
         foraging_30s,
         prop_foraging_30s,
         counted_when, cum_prop, 
         s_cum_prop, s2_cum_prop)

## Load the foraging data for 2021 dataset
filename3 <- "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data/data/foraging_data.csv"

tidy.for.2021 <-
  read.csv(filename3) %>% as_tibble() %>%
    mutate(date = as.Date(date, format="%m/%d/%y")) %>%
    mutate(colonyID = as.factor(colonyID)) 
  
tidy.for.2021 <- 
  tidy.for.2021 %>% 
  ## identify max foragers for each colony
  group_by(colonyID) %>% 
  summarise(max_foraging_30s = max(foraging_30s)) %>% 
  ungroup() %>% 
  right_join(tidy.for.2021) %>% 
  mutate(prop_foraging_30s = foraging_30s/max_foraging_30s) %>% 
  mutate_if(is.numeric, round, 2)

s.tidydat.2021 <- 
  old.summ %>%
  mutate(colonyID = as.factor(colonyID)) %>% 
  mutate(obs_day = parse_number(as.character(counted_when))) %>% 
  filter(obs_day==1) %>% 
  left_join(tidy.for.2021, by=c("colonyID", "date")) %>% 
  select(colonyID:bead_color, obs_day, foraging_30s, prop_foraging_30s) %>% 
  
  mutate(colonyID = as.double(colonyID)) %>%
  right_join(old.summ, by=c("colonyID", "date", "bead_color")) %>% 
  
  group_by(colonyID,bead_color) %>%  
  mutate(foraging_30s = foraging_30s[1]) %>% 
  mutate(prop_foraging_30s = prop_foraging_30s[1]) %>% 
  mutate(obs_day = paste0("D",obs_day[1])) %>% 
  ungroup() %>% 
  
  mutate(s_cum_prop = round(cum_prop/foraging_30s, 2)) %>%
  mutate(s2_cum_prop = round(cum_prop/prop_foraging_30s, 2)) %>%
  
  na.omit() %>% 
  # filter(s_cum_prop <= 1) %>%
  
  select(colonyID:bead_color,
         age_2022:age_class, 
         beads_collected,
         foraging_30s,
         prop_foraging_30s,
         counted_when, cum_prop, 
         s_cum_prop, s2_cum_prop)


## combine the two files
s.tidydat <- 
  rbind(s.tidydat.2021, s.tidydat.2022) %>% 
    mutate(counted_when = factor(counted_when,
                               levels = c("D1_bN", "D1_aN",
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  ## 4 age-classes
  # mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
  #                           ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
  #                                  ifelse(age_obs >=5 & age_obs < 10, "5-9_yr",
  #                                                "10+_yr")))) %>% 
  # mutate(age_class = factor(age_class, 
  #                           levels = c("1-2_yr", "3-4_yr",
  #                                      "5-9_yr", "10+_yr")))
  ## 3 age-classes
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
                            ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
                                                 "5+_yr"))) %>%
  mutate(age_class = factor(age_class, levels = c("1-2_yr", "3-4_yr","5+_yr")))

## order the colony IDs according to age
s.tidydat <- s.tidydat %>% 
  mutate(colonyID = factor(colonyID, levels = which.colonies %>% arrange(age_obs) %>% pull(1) %>% unique()))

## Data cleanup prior to running models
#### NEED TO BE FIXED ##
## ISSUE 1
# There are multiple foraging data entries for colony 1536, that needs to be corrected
# For now, let's look at all colonies, except 1536
s.tidydat <- s.tidydat %>% filter(colonyID != "1536")
## ISSUE 2
# There are two colonies for which we have collected more than 100% of the beads collected
# For now, let's remove those colonies
s.tidydat <- s.tidydat %>% 
  filter(!(colonyID == "1533" & bead_color == "green")) %>% 
  filter(!(colonyID == "1440" & bead_color == "blue"))
## ISSUE 3
# There are colonies that have no foraging data (or the colony did not forage) on D1 of bead experiments, 
# causing Inf values in s_cum_prop and s2_cum_prop
# For now, let's remove those colony-experiment-days
s.tidydat <- 
  s.tidydat %>% 
  # filter(is.infinite(s_cum_prop) | is.infinite(s2_cum_prop)) %>% 
  filter(!(colonyID == "1570" & bead_color == "blue")) %>% 
  filter(!(colonyID == "1536" & bead_color == "blue"))

```


### Compare foraging across age-class

### Fig. 5-7

```{r plot_foraging_ageclass}
## Plot 1
bd.saveplot(
  name = "05figure_foragingbeadcolonies_v1",
  width = 20, height = 10,
  s.tidydat %>% 
  select(colonyID, bead_color, age_obs, age_class, foraging_30s) %>% 
  distinct() %>%
  ggplot(aes(x=as.factor(age_obs), y=foraging_30s)) +
  geom_boxplot(alpha=0.7, outlier.color = NA) +
    geom_jitter(width = 0.1, size=3, alpha=0.6) +
    theme_Publication(base_size = 13) +
    labs(title = "Foraging (in+out) | bead colonies",
       subtitle = "observations made on Day1 of bead expmts.",
       y = "foraging (ants per 30s)",
       x = "age at observation") +
    theme(plot.subtitle=element_text(hjust=0.5, size=10),
          legend.title = element_blank()) +
    theme(legend.position="none")
)

## Plot 2
bd.saveplot(
  name = "06figure_foragingbeadcolonies_v2",
  width = 20, height = 10,
  tidy.for %>% 
  select(date, colonyID=colony_ID, foraging_30s) %>% 
  right_join(which.colonies, by="colonyID") %>% 
  ggplot(aes(x=as.factor(age_obs), y=foraging_30s)) +
  geom_boxplot(alpha=0.7) +
  geom_jitter(width = 0.1, size=2, alpha=0.6) +
  theme_Publication(base_size = 13) +
    labs(title = "Foraging (in+out) | bead colonies",
       subtitle = "observations made on bead colonies on all days",
       y = "foraging (ants per 30s)",
       x = "age at observation") +
    theme(plot.subtitle=element_text(hjust=0.5, size=10),
          legend.title = element_blank()) +
    theme(legend.position="none")
    # ## add text labels
    # ggrepel::geom_label_repel(data = . %>% 
    #                 mutate(label = ifelse(foraging_30s > 100,
    #                                       colonyID, "")) %>% 
    #                 group_by(colonyID) %>% 
    #                 mutate_at(vars(label), funs(replace(., duplicated(.), ""))),
    #                 force = 11,
    #                 aes(label = label), 
    #                 box.padding = 0.3,
    #                 show.legend = FALSE)
)

## Plot 3
bd.saveplot(
  name = "07figure_foragingallcolonies",
  width = 25, height = 10,
  tidy.for.2022 %>% 
  mutate(colony_ID = as.numeric(as.character(colony_ID))) %>% 
  right_join(colony.dat %>% select(colony_ID = colonyID, age_2022)) %>% 
  na.omit() %>% 
  mutate(age_obs = age_2022) %>% 
  ggplot(aes(x=as.factor(age_obs), y=foraging_30s)) +
  # geom_jitter(width = 0.2, size=0.5, alpha=0.2) +
  geom_boxplot(alpha=0.5, fill="#C5E4BA") +
    theme_Publication(base_size = 15) +
    labs(title = "Foraging (in+out) | ALL colonies",
       subtitle = "observations made on all colonies in 2022 and bead colonies in 2021",
       y = "foraging (ants per 30s) \n(mean ± 95% CI)",
       x = "age at observation") +
    # ylim(c(0,200)) +
    theme(plot.subtitle=element_text(hjust=0.5, size=10),
          plot.caption = element_text(hjust=1, size=10),
          legend.title = element_blank()) +
    theme(legend.position="none")
)

```


### Plot normalized bead returns

### Fig. 8-9

```{r plot_normalized_bead_returns}
## All colonies - 1
bd.saveplot(
  name="08figure_beadall_standardized_v1",
  width = 25, height = 20,
  ##
  s.tidydat %>% 
  ggplot(aes(x=counted_when, y=s_cum_prop, fill=age_class, color=age_class, group=age_class)) +
  # geom_jitter(size=2, alpha=0.4, width = 0.2) +
  geom_line(aes(group=interaction(colonyID,bead_color)),
            size=2, alpha=0.7) +
  facet_grid(age_class~bead_color) +
  # coord_flip() +
  scale_y_continuous(n.breaks = 3, labels = c("","0.05","0.10")) +
  scale_x_discrete(labels = c("D1_bN", "", "", "D2_bN", "", "D3_bN", "", "")) +
  theme_Publication(base_size = 20) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1),
        legend.position = "none")+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       caption = "\nbN = before NOON, D1 = Day 1")
)

```

### Plot normalized bead returns v2

> Standardizing by proportion of foragers outside on Day 1 of bead experiments

The idea is to divide the cumulative proportion of beads returned by the proportion of foragers (among the max foragers that were observed for a given colony) out on Day 1 of the bead experiments 

```{r standardizing_v2}
bd.saveplot(
  name="09figure_beadall_standardized_v2",
  width = 25, height = 20,
  ##
  s.tidydat %>% 
  ggplot(aes(x=counted_when, y=s2_cum_prop, fill=age_class, color=age_class, group=age_class)) +
  # geom_jitter(size=2, alpha=0.4, width = 0.2) +
  geom_line(aes(group=interaction(colonyID,bead_color)),
            size=2, alpha=0.7) +
  facet_grid(age_class~bead_color) +
  # coord_flip() +
  scale_y_continuous(n.breaks = 3, labels = c("","2.5","5.0")) +
  scale_x_discrete(labels = c("D1_bN", "", "", "D2_bN", "", "D3_bN", "", "")) +
  theme_Publication(base_size = 20) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1),
        legend.position = "none")+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       caption = "\nbN = before NOON, D1 = Day 1")
)
```

___

## Prep data for modeling

```{r prep_mod_data}
m <- 
  s.tidydat %>% distinct() %>% filter(cum_prop>1) %>% select(colonyID, bead_color) %>% 
  distinct() %>% 
  as.matrix()

s.tidydat.mod <- s.tidydat %>% 
  # remove colony-trials that have a cumulative proportion above 1
  filter(!(colonyID%in%m[,1] & bead_color%in%m[,2]))


# Make the prop_foraging a group (high/low/medium)
quartiles.prop_foraging <- 
  s.tidydat.mod %>% 
  pull(prop_foraging_30s) %>% 
  unique() %>% 
  summary() %>% 
  as.vector()
q1.prop_foraging <- quartiles.prop_foraging[2]
q3.prop_foraging <- quartiles.prop_foraging[5]
s.tidydat.mod <- 
  s.tidydat.mod %>% 
  mutate(foraging_class = ifelse(prop_foraging_30s <= q1.prop_foraging, "low",
                                 ifelse(prop_foraging_30s >= q3.prop_foraging, "high", "average"))) %>% 
  mutate(foraging_class = factor(foraging_class, levels = c("low","average","high"))) %>% 
  select(colonyID:bead_color, 
         beads_collected, foraging_class, 
         counted_when:cum_prop, everything())

## Summarize the data
s.tidydat.mod %>% 
  group_by(age_class, foraging_class) %>% 
  summarize(n.colonies = length(unique(colonyID)),
            n.replicates = length(unique(bead_color)),
            # which.colonies = paste(unique(colonyID),collapse = ", "),
            which.reps = paste(unique(bead_color),collapse = ", "))

## Convert counted-when into a numeric vector that suggests time passed
M <- 8; bN <- 11; aN <- 14;
s.tidydat.mod2 <-
  s.tidydat.mod %>% 
  mutate(c2 = counted_when) %>% 
  separate(c2, c("day", "time"), "_") %>% 
  mutate(t2 = ifelse(time == "M", M,
                     ifelse(time == "bN", bN, aN))) %>% 
  mutate(d2 = readr::parse_number(day)) %>% 
  mutate(t.since = (d2-1)*24+(t2-M)) %>% 
  select(colonyID:bead_color, 
         beads_collected, foraging_class, 
         day, t2, t.since, everything())

```


## EDA

## Let's plot hypothesized relations

### Fig. 10-13

```{r plot_modeled_relations}
# 1. (Cum-prop v. time) v. age-class
bd.saveplot(
  name="10figure_beadreturns_v_time",
  width = 30, height = 20,
  s.tidydat.mod2 %>% 
  ggplot(aes(y=cum_prop,x=t.since)) +
  geom_jitter(width=0.1) +
  geom_smooth(method = "lm", se = FALSE, col="red") +
  facet_grid(age_class~bead_color) +
  labs(x="time since beads were presented (h)",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")
)


# 2. Cum-prop v. foraging
bd.saveplot(
  name="11figure_beadreturns_v_time",
  width = 30, height = 20,
  s.tidydat.mod2 %>% 
  ggplot(aes(y=cum_prop,x=t.since)) +
  geom_jitter(width=0.1) +
  geom_smooth(method = "lm", se = FALSE, col="red") +
  facet_grid(age_class~foraging_class) +
  labs(x="time since beads were presented (h)",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")
)


# 3. Cum-prop v. age-class
bd.saveplot(
  name="12figure_beadreturns_v_time",
  width = 35, height = 20,
  s.tidydat.mod2 %>% 
  ggplot(aes(y=cum_prop,x=age_class)) +
  geom_boxplot(alpha=0.6, outlier.color = NA) +
  geom_jitter(width=0.01, alpha=0.3) +
  facet_grid(foraging_class~counted_when) +
  labs(x="age-class",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle=90))
)
bd.saveplot(
  name="13figure_beadreturns_v_time_v2",
  width = 30, height = 20,
  s.tidydat.mod2 %>% 
    filter(time=="bN" & day!="D4") %>% 
  ggplot(aes(y=cum_prop,x=age_class)) +
  geom_boxplot(alpha=0.6, outlier.color = NA) +
  geom_jitter(width=0.1, alpha=0.3, size=2) +
  facet_grid(foraging_class~counted_when) +
  labs(x="age-class",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle=90))
)

```


### Fig. 14-16

```{r plot_modeled_relations}
## create a subset of the data with all observations for day 1 and 2
s.tidydat.mod3 <- 
  s.tidydat.mod2 %>% 
    filter(counted_when %in% c("D1_bN", "D2_M", "D2_bN", "D2_aN"))

# 1. (Cum-prop v. time) v. age-class
bd.saveplot(
  name="14figure_subset_beadreturns_v_time",
  width = 30, height = 20,
  s.tidydat.mod3 %>% 
  ggplot(aes(y=cum_prop,x=t.since)) +
  geom_jitter(width=0.1) +
  geom_smooth(method = "lm", se = FALSE, col="red") +
  facet_grid(age_class~bead_color) +
  labs(x="time since beads were presented (h)",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")
)


# 2. Cum-prop v. foraging
bd.saveplot(
  name="15figure_subset_beadreturns_v_time",
  width = 30, height = 20,
  s.tidydat.mod3 %>% 
  ggplot(aes(y=cum_prop,x=t.since)) +
  geom_jitter(width=0.1) +
  geom_smooth(method = "lm", se = FALSE, col="red") +
  facet_grid(age_class~foraging_class) +
  labs(x="time since beads were presented (h)",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none")
)


# 3. Cum-prop v. age-class
bd.saveplot(
  name="16figure_subset_beadreturns_v_time",
  width = 35, height = 20,
  s.tidydat.mod3 %>% 
  ggplot(aes(y=cum_prop,x=age_class)) +
  geom_boxplot(alpha=0.6, outlier.color = NA) +
  geom_jitter(width=0.01, alpha=0.3) +
  facet_grid(foraging_class~counted_when) +
  labs(x="age-class",
       y="cumulative proportion\nof beads returned") +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c("","0.5","1.0")) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle=90))
)
```


## lme4::lmer()

```{r prelim_analyses_v1}
pacman::p_load(lme4)
mod.full.1a <- lme4::glmer(cum_prop ~ age_class*t.since + foraging_class + (1|bead_color) + (1|colonyID),
                 data=s.tidydat.mod2,
                 control = lme4::glmerControl(optimizer = "bobyqa"),
                 family = binomial(link = "logit"))
# mod.full.1b <- lme4::glmer(cum_prop ~ age_class*t.since + foraging_class + (1|bead_color),
#                data=s.tidydat.mod2,
#                family = binomial(link = "logit"))
# mod.full.1c <- lme4::glmer(cum_prop ~ age_class*t.since*foraging_class + (1|colonyID),
#                data=s.tidydat.mod2,
#                family = binomial(link = "logit"))
# mod.full.1d <- lme4::glmer(cum_prop ~ age_class*t.since + foraging_class + (1|colonyID:bead_color) + (1|colonyID),
#                  data=s.tidydat.mod2,
#                  family = binomial(link = "logit"))
# bbmle::AICctab(mod.full.1a, mod.full.1b, mod.full.1c, mod.full.1d)
# Here are the AIC values:
# (1a = 348, 1b = 359, 1c = 343, 2 = 356, 3 = 351)
#
## Using prop_foraging_30s instead of foraging-class
mod.full.2 <- lme4::glmer(cum_prop ~ age_class*t.since + foraging_class + foraging_class:age_class + (1|bead_color) + (1|colonyID),
                           family = binomial(link = "logit"),
                          control = lme4::glmerControl(optimizer = "bobyqa"),
                 data=s.tidydat.mod2)
bbmle::AICctab(mod.full.1a,mod.full.2)
#
### Conclusion: The winner is mod.full.1a !!!

## Let's run diagnostic on the model
plot(mod.full.1a)
## as per Zuur "beginneers guide to GAMs with R"
mod.validate(mod.full.1a)

# mod.red.1a
# no interaction between age-class and counted-when
# 
mod.red.1a <- lme4::glmer(cum_prop ~ age_class + t.since + foraging_class + (1|bead_color) + (1|colonyID),
                 data=s.tidydat.mod2,
                 control = lme4::glmerControl(optimizer = "bobyqa"),
                 family = binomial(link = "logit"))
anova(mod.full.1a,  mod.red.1a, test="Chisq")
#
### Ans: Yes, there is a significant interaction.

# mod.red.1b
# 
mod.red.1b <- lme4::glmer(cum_prop ~ age_class*t.since+ (1|bead_color) + (1|colonyID),
                 data=s.tidydat.mod2,
                 control = lme4::glmerControl(optimizer = "bobyqa"),
                 family = binomial(link = "logit"))
anova(mod.full.1a,  mod.red.1b, test="Chisq")
#
### Ans: Yes, colony foraging is significantly correlated with colony's trash removal

## Get a tidy summary table
broom::tidy(mod.full.1a)

## Source: https://stats.oarc.ucla.edu/r/dae/mixed-effects-logistic-regression/
# m <- mod.full.1a
# print(m, corr=F)
# 
# se <- sqrt(diag(vcov(m)))
# # table of estimates with 95% CI
# (tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *
#     se))
# # If we wanted odds ratios instead of coefficients on the logit scale, we could exponentiate the estimates and CIs
# exp(tab)
# 
# lattice::dotplot(ranef(m, which = "colonyID", condVar = TRUE), scales = list(y = list(alternating = 0)))
# lattice::dotplot(ranef(m, which = "bead_color", condVar = TRUE), scales = list(y = list(alternating = 0)))


# Source: http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html?fbclid=IwAR3q1t3PGkPT3Q4yuIaEpV6ehqfKB65IMyLlqH4JWqJnKluFb6xnhWDVmQM#beta-glmms
# summary(lme4::allFit(mod.full.1a))

# ## Calculate the p-values for the fixed effects
# ## Source: https://mspeekenbrink.github.io/sdam-r-companion/linear-mixed-effects-models.html
afex::mixed(formula(mod.full.1a), data=s.tidydat.mod2,
            check_contrasts = T, test_intercept = F, method="S")

```


## glmmTMB::glmmTMB()

> Sources: 

[Tutorial by Ben Bolker](https://mirrors.pku.edu.cn/CRAN/web/packages/glmmTMB/vignettes/glmmTMB.pdf)

```{r glmm_tmb}

tmb.full.1a <- glmmTMB::glmmTMB(cum_prop ~ age_class*t.since + foraging_class + (1|bead_color) + (1|colonyID),
                 data=s.tidydat.mod2,
                 family = binomial(link = "logit"))

bbmle::AICctab(mod.full.1a, tmb.full.1a)
```

So, it appears that the logistic regression model we build using the lme4::glmer function performs better that the glmmTMB one.


> Other possibilities of modeling frameword 

One other possibility that is fairly common in ecoloy is the use of Beta regression using the betareg() in betareg package. However, I am not sure if we can add random effects in the model.



___

## Part TWO

## Foraging v. bead returns

### foraging on Day 1

```{r for_v1}
s.tidydat.mod %>% 
  group_by(colonyID, bead_color) %>% 
  summarize(mBeadreturned = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  
  select(colonyID:age_class, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  
  ggplot(aes(x=foraging_30s, y=mBeadreturned)) +
    geom_jitter(aes(fill=age_class), size=5, shape=21, width=0, alpha=0.8) +
    facet_wrap(age_class~., nrow = 2) +
    geom_smooth(method = "gam") +
    scale_color_manual(values=age_colors) +
    # scale_x_continuous(limits = c(0,1)) +
    theme_Publication(base_size = 15) +
  labs(x = "total foraging for 30s on D1",
       y= "proportion of beads \nreturned at end of exp.",
       subtitle = "",
       caption = "")

## 2
s.tidydat.mod %>% 
  group_by(colonyID, bead_color) %>% 
  summarize(mBeadreturned = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  
  select(colonyID:age_class, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  
  ggplot(aes(x=foraging_30s, y=mBeadreturned, col=age_class)) +
  geom_hline(yintercept=0.75, col=age_colors[4], size=1) +
  geom_vline(xintercept=100, col=age_colors[4], size=1) +
  # geom_smooth(method = "lm", se=F, size=1.5, alpha=0.7) +
  geom_point(size=3, alpha=0.8, width = 0.2) +
  # geom_line(aes(group=colonyID), size=2, alpha=0.7) +
  facet_wrap(colonyID~., nrow=6) +
  theme_bw(base_size = 8) +
  # theme(axis.title.x=element_blank(),
  #       axis.text.x=element_blank())+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "total foraging for 30s on D1",
       y= "proportion of beads \nreturned at end of exp.",
       subtitle = "",
       caption = "")


```


### proportion of foragers on Day 1

```{r for_v2}
## 1
s.tidydat.mod %>% 
  group_by(colonyID, bead_color) %>% 
  summarize(mBeadreturned = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  ungroup() %>% 
  select(colonyID:age_class, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  
  ggplot(aes(x=prop_foraging_30s, y=mBeadreturned)) +
    geom_jitter(aes(fill=age_class), shape=21, width=0, alpha=0.8, size=5) +
    facet_wrap(age_class~., nrow = 2) +
    geom_smooth(method = "gam") +
    scale_color_manual(values=age_colors) +
    # scale_x_continuous(limits = c(0,1)) +
    theme_Publication(base_size = 15) +
  labs(x = "proportion of foragers out on D1",
       y= "proportion of beads \nreturned at end of exp.",
       subtitle = "",
       caption = "")

## 2
s.tidydat.mod %>% 
  group_by(colonyID, bead_color) %>% 
  summarize(mBeadreturned = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  ungroup() %>% 
  select(colonyID:age_class, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  
  ggplot(aes(x=prop_foraging_30s, y=mBeadreturned, col=age_class)) +
  geom_hline(yintercept=0.75, col=age_colors[4], size=1) +
  geom_vline(xintercept=0.75, col=age_colors[4], size=1) +
  # geom_smooth(method = "lm", se=F, size=1.5, alpha=0.7) +
  geom_point(size=3, alpha=0.8, width = 0.2) +
  # geom_line(aes(group=colonyID), size=2, alpha=0.7) +
  facet_wrap(colonyID~., nrow=6) +
  theme_bw(base_size = 8) +
  # theme(axis.title.x=element_blank(),
  #       axis.text.x=element_blank())+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  labs(x = "proportion of foragers out on D1",
       y= "proportion of beads \nreturned at end of exp.",
       subtitle = "",
       caption = "")
```

### foraging and bead returns on Day 1

```{r for_v3}
## 1
s.tidydat.mod %>% 
  filter(counted_when %in% c("D1_bN", "D1_aN")) %>% 
  group_by(colonyID, bead_color) %>% 
  summarize(mBeadreturned = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  ungroup() %>% 
  select(colonyID:age_class, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  
  ggplot(aes(x=prop_foraging_30s, y=mBeadreturned)) +
    geom_jitter(aes(fill=age_class), shape=21, width=0, alpha=0.8, size=5) +
    facet_wrap(age_class~., nrow = 1) +
    geom_smooth(method = "gam") +
    scale_color_manual(values=age_colors) +
    # scale_x_continuous(limits = c(0,1)) +
    theme_Publication(base_size = 15) +
    labs(x = "proportion of foragers out on D1",
       y= "max. proportion \nof beads returned on Day 1",
       subtitle = "",
       caption = "")

## 2
s.tidydat.mod %>% 
  filter(counted_when %in% c("D1_bN", "D1_aN")) %>% 
  group_by(colonyID, bead_color) %>% 
  summarize(mBeadreturned = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  ungroup() %>% 
  select(colonyID:age_class, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  
  ggplot(aes(x=prop_foraging_30s, y=mBeadreturned, col=age_class)) +
  geom_hline(yintercept=0.75, col=age_colors[4], size=1) +
  geom_vline(xintercept=0.75, col=age_colors[4], size=1) +
  # geom_smooth(method = "lm", se=F, size=1.5, alpha=0.7) +
  geom_point(size=3, alpha=0.8, shape=21, aes(fill=age_class), col="black") +
  # geom_line(aes(group=colonyID), size=2, alpha=0.7) +
  facet_wrap(colonyID~., ncol=6) +
  theme_bw(base_size = 8) +
  # theme(axis.title.x=element_blank(),
  #       axis.text.x=element_blank())+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "proportion of foragers out on D1",
       y= "max. proportion \nof beads returned on Day 1",
       subtitle = "",
       caption = "")
```

### relative allocation in foraging and nest maintenance

```{r for_v4}
## First, let's check how many observations we have for each time point of observation.
# s.tidydat.mod %>% 
#   group_by(bead_color, age_class, counted_when) %>% 
#   summarise(num = n()) %>% 
#   # left_join(s.tidydat.mod %>% select(colonyID, age_class), by=) %>% 
#   ggplot(aes(x=counted_when, y=num)) +
#   geom_jitter(aes(group=bead_color, col=bead_color), size=6, width = 0.2) +
#   facet_grid(~age_class) +
#   # scale_color_manual(values=age_colors) +
#   theme_Publication(base_size = 20) +
#   scale_y_continuous(limits = c(0,25))
#### So, we have good and bad news:
####  we do have D2_M data for most bead-colonies in 2022, but not for 2021

## 1
s.tidydat.mod %>% 
  filter(counted_when %in% c("D2_M")) %>% 
  group_by(colonyID) %>% 
  summarize(max_D2M = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  ungroup() %>% 
  select(colonyID:age_class, cum_prop, counted_when, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  na.omit() %>% 
  
  filter(counted_when %in% c("D2_M")) %>% 
  mutate(prop_nm_D2M = round(cum_prop/max_D2M,2)) %>% 
  
  ggplot(aes(x=prop_foraging_30s, y=prop_nm_D2M)) +
    geom_jitter(aes(fill=age_class), shape=21, width=0, alpha=0.8, size=5) +
    facet_wrap(age_class~., nrow = 1) +
    geom_smooth(method = "gam") +
    scale_color_manual(values=age_colors) +
    # scale_x_continuous(limits = c(0,1)) +
    theme_Publication(base_size = 15) +
    labs(x = "proportion of foragers out on Day 1 (divided by max observed)",
       y= "proportion of beads returned on Day 1 \n(divided by max observed on Day1)",
       subtitle = "",
       caption = "")

## 2
s.tidydat.mod %>% 
  filter(counted_when %in% c("D2_M")) %>% 
  group_by(colonyID) %>% 
  summarize(max_D2M = max(cum_prop)) %>% 
  right_join(s.tidydat.mod) %>% 
  ungroup() %>% 
  select(colonyID:age_class, cum_prop, counted_when, foraging_30s, prop_foraging_30s, -date) %>% 
  distinct() %>% 
  na.omit() %>% 
  
  filter(counted_when %in% c("D2_M")) %>% 
  mutate(prop_nm_D2M = round(cum_prop/max_D2M,2)) %>% 
  
  ggplot(aes(x=prop_foraging_30s, y=prop_nm_D2M, col=age_class)) +
  geom_hline(yintercept=0.5, col=age_colors[4], size=1) +
  geom_vline(xintercept=0.5, col=age_colors[4], size=1) +
  # geom_smooth(method = "lm", se=F, size=1.5, alpha=0.7) +
  geom_point(size=3, alpha=0.9, aes(fill=age_class, shape=bead_color)) +
  facet_wrap(colonyID~., ncol=6) +
  theme_bw(base_size = 8) +
  scale_y_continuous(limits = c(0,1.1), breaks = c(0,0.5,1.0)) +
  # theme(axis.title.x=element_blank(),
  #       axis.text.x=element_blank())+
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "proportion of foragers out on Day 1 (divided by max observed)",
       y= "proportion of beads returned on Day 1 \n(divided by max observed on Day1)",
       subtitle = "",
       caption = "")
```

___

## Extras

> How many colonies, and which colonies, do we have for each age-class? 

```{r tallying_22Aug22_green_beads}
s.tidydat %>% 
  mutate(age_obs = factor(paste0(age_obs, " yr"), 
                          level=paste0(sort(unique(.$age_obs)), " yr"))) %>%
  group_by(age_obs) %>% 
  summarize(n_colonies = length(unique(colonyID)),
            which_colonies = paste(as.character(unique(colonyID)), collapse = ", ")) %>% 
  
  kableExtra::kable(caption = "Colonies used in bead experiments", align = "c") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
  

```


