---
title: "Deborah_bead_data"
author: "Biplabendu Das"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    keep_md: no
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

set.seed(420)
rm(list = ls())

```


```{r housekeeping, include=FALSE}
## Load the libraries
pacman::p_load(tidyverse, viridis, conflicted)
pacman::p_load(ggplot2, patchwork)
# pacman::p_load(RSQLite, dbplyr)

## set conflict preference
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("union", "dplyr")
conflict_prefer("lmer", "lme4")

## set path to your working directory
path_to_repo = "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data"

## load functions
# customized theme for publication quality figures
source(paste0(path_to_repo,"/functions/theme_publication.R"))
# # function to perform enrichment analysis
# source(paste0(path_to_repo,"/functions/enrichment_analysis.R"))
# # # function to plot z-scores (Cflo genes only)
# source(paste0(path_to_repo,"/functions/plot_zscores.R"))

```


```{r set_parameters}
## Set color scheme for age classes
age_colors <- c("#8D72A1", "#F26C7F", "#FBAD73", "#9CD5A2", "#F29544")
```

## Overview/Goals

Email from Deborah: 

This data set comes from an experiment with harvester ants last summer. Harvester ant foragers bring in lots of seeds and parts of plants, some of which are rejected by ants inside the nest. Eventually, the refuse is carried out and put on the nest mound midden. 

> Here we presented small plastic beads coated in oil which the ants brought into the nest, and were later put back out on the midden, outside the colony, close to the nest entrance.

## Bead data 2021

```{r notes_v1, eval=F, include=F, echo=F}
# ## Part 1: INSPECT DATA
# 
# Two data sheets are available in the Excel workbook. For convenience of analysis, the excel sheets are saved individually as csv files and, if necessary, the column names for each were modified.
# 
# Sheet 1: foraging_data.csv
# - contains three columns: date, colonyID, foraging_30s
# 
# Sheet 2: beads_out.csv
# - contains XXX cols: num_out, date, colonyID, time_put_out, n_collected, n_returned_DayX (X %in% 1:6)
# 
# Brief description of the columns:
# - date = date of observation
# - colonyID = unique colony identifier
# - time_put_out = time-of-day at which beads were put out
# - n_collected = number of beads that had been brought into the nest at the end of foraging on Day 1
# - n_returned_DayX = number of beads that had been brought *out of the nest* by the end of **Day X**
# 
# Note; 
# NA means the colony is not active that day
# NM means that the colony was doing nest maintenance, when they carry out refuse including beads
# 
# Some additional info:
# 
# Beads coated with oil were put out in 6 colonies on day 1, color green, then another 6 colonies on day 2, color green, then again with first 6 on day 3 color pink, then again with 2nd 6 on day 4 color pink, back to green days 5 and 6
# 
# Numbers of beads brought out are listed separately by color. E.g., colony 1131 took out 18 green beads on day 2 of green 8-24, and 3 pink beads on day 2 of pink 8-26.
# 
# For the 3rd round, for whatever number of green beads from the first round were not out by 8-27 or 8-28, then we don't know which day those were from.

```



### Load data

```{r load_data}
# read foraging_data.csv
for.dat <- read.csv(paste0(path_to_repo,"/data/foraging_data.csv"),
                    stringsAsFactors = F, header = T) %>% as_tibble()

# read bead data
bead.dat <- read.csv(paste0(path_to_repo, "/data/beads_out_new.csv"),
                stringsAsFactors = F, header = T) %>% as_tibble()

# read colony-age data
colony.dat <- read.csv(paste0(path_to_repo, "/data/foraging_counts_08_13_22.csv"),
                       stringsAsFactors = F, header = T) %>% as_tibble() %>% 
  select(colonyID = colony_ID, everything()) %>% 
  # mutate(colonyID = as.factor(colonyID)) %>% 
  mutate(age_class = ifelse(age_2022 <= 2, "1-2_yr_old",
                                  ifelse(age_2022 > 2 & age_2022 <= 4, "3-4_yr_old",
                                         "older_than_5_yr")))
```


### Summarize the data

```{r summarize_data}

writeLines("What are the names of the colonies? \nShown as: Colony-ID")
bead.dat %>% 
  select(colonyID) %>% 
  distinct() %>% 
  left_join(colony.dat, by="colonyID") %>%
  head() %>% 
  arrange(age_2022)


# Specify the dates for green/pink beads
green.days <- c("8/23/21","8/24/21", "8/27/21","8/28/21") %>% as.Date(format = "%m/%d/%y")
pink.days <- c("8/25/21","8/26/21") %>% as.Date(format = "%m/%d/%y")
```

### Clean Data

```{r clean_1}
df <-
  bead.dat %>%
  
  # retain only useful data
  na.omit() %>% 
  
  # arrange by colony IDs
  arrange(colonyID) %>% 
  
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
  mutate(colonyID = factor(colonyID)) %>% 
  
  # add a column that specifies the color of beads used
  mutate(bead_col = ifelse(date %in% green.days[1:2], "green",
                           ifelse(date %in% green.days[3:4], "green-2",
                           ifelse(date %in% pink.days, "pink", NA)))) %>%
  
  mutate(batch = ifelse(date %in% c(green.days[c(1,3)],pink.days[1]), "batch_A",
                        ifelse(date %in% c(green.days[c(2,4)],pink.days[2]), "batch_B",NA))) %>% 
  
  
  # # SIDE NOTE: Explore the original dataset
  # mutate_at(vars(colnames(.)[names(.) %in% c(paste0("n_returned_Day",1:6))]),
  #       .funs = funs(ifelse(.=="", NA, as.character(.)))) %>% 
  # mutate_at(vars(colnames(.)[names(.) %in% c(paste0("n_returned_Day",1:6))]),
  #       .funs = as.numeric) %>% head()
  
  
  # CLEAN: Keep only data for first three days of observations after presenting beads
  select(num_out:n_returned_Day3, bead_col, batch) %>%
  
  # format columns appropriately
  mutate(n_collected = as.numeric(readr::parse_number(n_collected))) %>% 
  mutate(n_returned_Day1 = as.numeric(n_returned_Day1)) %>% 
  mutate(n_returned_Day2 = as.numeric(n_returned_Day2)) %>% 
  mutate(n_returned_Day3 = as.numeric(n_returned_Day3))
  

# query 1  
# df %>% 
#   # summarize the data by batch and bead_col
#   group_by(batch, bead_col, date) %>%
#   summarize(n_colonies = n())
  
# query 2  
# It seems the data for 25-Aug-2021, Colony-1474 is missing (or too little);  
df <- 
  df %>% 
  # remove it
  filter(!is.na(n_collected))

```


CLEAN UP:

1. There are four colonies (1 green, 3 pink) that collect <30 of the 50 beads presented. [n_collected = 8, 9, 17, 20]. It might be a good idea to leave them out before we analyze the dat, at least for Colonies that collected ≤10 beads of the 50 presented. [cleanup 1]

2. Data for Colony 1288 (green-1; 24 to 26 Aug) shows that 48 beads were presented but 50 beads were collected at the refuse pile. Removing that from our dataset as I am unsure of the source of the error. [cleanup 2]


```{r remove_sample}
dat <-
  df %>%
  # remove the time column (not necessary; almost the same time) 
  select(-4) %>% 
  filter(!(bead_col=="pink" & colonyID %in% c("969","1338"))) %>% # cleanup 1
  filter(!(bead_col=="green" & colonyID == "1288")) # cleanup 

writeLines("Which colony-days remain after cleanup?")
dat %>% 
  select(date, colonyID) %>% 
  distinct() %>% 
  mutate(colonyID = as.numeric(as.character(colonyID))) %>%
  left_join(colony.dat, by="colonyID") %>% 
  group_by(colonyID, age_2022, age_class) %>% 
  summarize(n_days = n()) %>% 
  arrange(age_2022)
```


```{r prep_data}
## Part 5: Plot cum. prop.

tidydat.2021 <-
  dat %>% 
  
  na.omit() %>%
  
  ## only keep the colonies that collected more than 40 beads
  filter(n_collected >= 40) %>% 
  
  # Cumulative prop 
  # create the cumulative sums
  mutate(cum_prop_day1 = n_returned_Day1/n_collected,
         cum_prop_day2 = (n_returned_Day1+n_returned_Day2)/n_collected,
         cum_prop_day3 = (n_returned_Day1+n_returned_Day2+n_returned_Day3)/n_collected) %>% 
  
  select(num_out, date, colonyID, n_collected, bead_col, batch:cum_prop_day3) %>% 
  
  pivot_longer(cols = starts_with("cum_prop"),
               names_to = "day_of_obs",
               values_to = "cum_prop_returned") %>% 
  
  mutate(day_of_obs = as.numeric(readr::parse_number(day_of_obs))) %>% 
  mutate(date = date+day_of_obs-1)
  
# # sumarrize the tidydat
# tidydat.2021 %>% 
#   group_by(batch, bead_col, date) %>%
#   summarize(n_colonies = n())
# 
# # what proportion of collected beads do they put out by day 3?
# tidydat.2021 %>% 
#   filter(day_of_obs=="3") %>%
# # 
# #   pull(n_collected) %>% sd()/sqrt(51)
# #   sd()/sqrt(17) # 17 = length of cum_prop_returned
# 
#   group_by(bead_col, batch) %>% 
#   summarize(n_colonies=n(),mean_prop_returned=mean(cum_prop_returned))

```


___

```{r abc}
# Selecting old colonies

old.summ <- 
  tidydat.2021 %>% 
  mutate(colonyID = as.character(colonyID) %>% as.numeric()) %>% 
  left_join(colony.dat, by="colonyID") %>% 
  # select(date, colonyID, bead_col, day_of_obs, cum_prop_returned, age_2022) %>% 
  
  mutate(age_obs = age_2022-1) %>% 
  
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr_old",
                            ifelse(age_obs >= 5, "older_than_5_yr", "3-4_yr_old"))) %>% 
  mutate(age_class = factor(age_class, levels = c("1-2_yr_old", "3-4_yr_old", "older_than_5_yr"))) %>% 
  
  mutate(day_of_obs = paste0("D",day_of_obs,"_bN")) %>% 
  mutate(day_of_obs = factor(day_of_obs, levels = c("D1_bN", "D2_bN", "D3_bN"))) %>% 
  
  mutate(bead_col = factor(paste0(bead_col,"-21"))) %>% 
  mutate(ID = paste0(colonyID,"_",bead_col, "_", age_obs)) %>% 
  
  ## change column names to make them compatible to the 2022 dataset
  select(colonyID, date,
         bead_color = bead_col,
         age_2022, age_obs, age_class, 
         beads_collected = n_collected,
         counted_when = day_of_obs,
         cum_prop = cum_prop_returned)


old.summ %>% 
  as_tibble() %>% 
  filter(counted_when=="D3_bN") %>% 
  group_by(colonyID,age_class) %>% 
  summarize(n_reps = n()) %>% 
  right_join(old.summ %>% select(colonyID, age_obs) %>% distinct()) %>% 
  arrange(age_obs)

old.summ %>%   
   mutate(ID = paste0(colonyID,"_",bead_color, "_", age_obs)) %>% 
  
  ggplot() + 
  geom_line(aes(x=factor(counted_when), 
                y=cum_prop, group=ID, col=ID),
            size=2,
            alpha=0.6) +
  facet_wrap(~ID, nrow = 2) +
  theme_Publication(base_size = 10) +
  # scale_color_manual(values=c("darkgreen","darkred")) +
  
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1),
                     labels = c("0", "", "0.4","","0.8",""),
                     limits = c(0,1)) +
  
  xlab("Day of obs") +
  ylab("Cumulutative proportion \nof beads returned")


old.summ %>% 
    
    ggplot(aes(x=counted_when, y=cum_prop, 
               fill=age_class, color=age_class, group=age_class)) +
    geom_jitter(size=2, alpha=0.4, width=0.2) +
    # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
    # geom_boxplot(alpha=0.3) +
    geom_smooth(method = "lm", formula = y~log2(x+1), 
                level=0.95,
                alpha=0.4, size = 2,
                aes(color=age_class)) +
  # geom_smooth(method = "lm", formula = y~x, alpha=0.3, aes(color=age_class, lty=age_class)) +
  
  facet_grid(~age_class) +
  # coord_flip() +
  # scale_y_continuous(n.breaks = 3,
  #                    labels = c("","0.5", "1")) +
  
  theme_Publication(base_size = 15) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  # scale_fill_discrete() +
  # scale_color_discrete()

  # scale_fill_viridis_d(option="cividis") +
  # scale_color_viridis_d(option = "cividis")
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "Line indicates y ~ log2(x+1) + 95% CI")

```

___


## Bead data 2022


```{r data_2022}
filename <- "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data/data/bead_data_2022/all_bead_data_2022/2022_bead_data_master_replicate_one_two_three_29Aug22.csv"
dat.2022 <- read.csv(filename,
                     header = T, stringsAsFactors = F,
                     na.strings = c(""," ","did_not_count","NA",
                                    "did_not_check",
                                    "it_rained_could_not_check")) %>% 
  as_tibble() %>% 
  
  # select the relevant columns
  select(date, colonyID, age_class, age_2022, 
         bead_color, beads_put_out, beads_collected,
         beads_returned_Day1_beforenoon,
         beads_returned_Day1_afternoon,
         beads_returned_Day2_morning,
         beads_returned_Day2_beforenoon,
         beads_returned_Day2_afternoon,
         beads_returned_Day3_beforenoon,
         beads_returned_Day4_morning,
         beads_returned_Day4_beforenoon) %>% 
  
  # calculate the cumulative proportion of beads returned
  group_by(date, bead_color) %>% 
    
  # filter out the colonies for which we still need to check the data
  filter(!(colonyID=="1560" & bead_color=="green")) %>% 
  filter(!(colonyID=="1288" & bead_color=="green")) %>%
  filter(!(colonyID=="1546" & bead_color=="green")) %>%
  
  # convert all the bead count columns as numeric
  mutate_at(vars(matches("beads_")), as.numeric) %>% 
  
  ungroup()


## Create a tidy dataset and calculate the cumulative proportion of beads returned by colonies
tidydat.2022 <-
  dat.2022 %>% 
  
  # replace all the NAs with 0 to be able to calculate proportions
  mutate_at(vars(matches("beads_")), function(x) ifelse(is.na(x), 0, x)) %>%
  
  # day 1 - before noon
  mutate(cum_prop_returned_D1_beforenoon = round(beads_returned_Day1_beforenoon/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D1_afternoon = round(cum_prop_returned_D1_beforenoon+beads_returned_Day1_afternoon/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D2_morning = round(cum_prop_returned_D1_afternoon+beads_returned_Day2_morning/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D2_beforenoon = round(cum_prop_returned_D2_morning+beads_returned_Day2_beforenoon/beads_collected,2)) %>%
  mutate(cum_prop_returned_D2_afternoon = round(cum_prop_returned_D2_beforenoon+beads_returned_Day2_afternoon/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D3_beforenoon = round(cum_prop_returned_D2_beforenoon+beads_returned_Day3_beforenoon/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D4_morning = round(cum_prop_returned_D3_beforenoon+beads_returned_Day4_morning/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D4_beforenoon = round(cum_prop_returned_D4_morning+beads_returned_Day4_beforenoon/beads_collected,2)) %>% 
  
  mutate(beads_collected_group = ifelse(beads_collected < 22, "less_than_22", 
                                        ifelse(beads_collected > 40, 
                                               "greater_than_40", "between_22_and_40"))) %>%
  arrange(age_2022, beads_collected_group) %>% 
  
  select(colonyID,
         age_2022, age_class,
         beads_collected_group,
         bead_color,
         cum_prop_returned_D1_beforenoon:cum_prop_returned_D4_beforenoon,
         everything())

  
```


```{r prep_data_for_plot}
tidydat.2022 %<>% 
  
  ## keep only colonies that collected at least 40 (of 50) beads offered
  filter(beads_collected_group == "greater_than_40") %>% 
  
  pivot_longer(cols = starts_with("cum_prop"),
               names_to = "counted_when",
               values_to = "cum_prop") %>% 
  
  ## format the ordinal columns correctly
  mutate(counted_when = ifelse(counted_when == "cum_prop_returned_D1_beforenoon", "D1_bN",
                            ifelse(counted_when == "cum_prop_returned_D1_afternoon", "D1_aN",
                                   ifelse(counted_when == "cum_prop_returned_D2_morning", "D2_M",
                                          ifelse(counted_when == "cum_prop_returned_D2_beforenoon", "D2_bN",
                                            ifelse(counted_when == "cum_prop_returned_D2_afternoon", "D2_aN",
                                            ifelse(counted_when == "cum_prop_returned_D3_beforenoon", "D3_bN",
                                            ifelse(counted_when == "cum_prop_returned_D4_morning", "D4_M",
                                                 "D4_bN")))))))) %>%
  mutate(counted_when = factor(counted_when, 
                               levels = c("D1_bN", "D1_aN", 
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  mutate(bead_color = factor(ifelse(bead_color %in% c("light_blue", "dark_blue"), "blue", bead_color),
                             levels = c("pink","green","blue")))  %>% 
  mutate(colonyID = factor(colonyID)) %>% 
  mutate(date = as.Date(date, format="%m/%d/%y")) %>% 
  
  mutate(age_obs = age_2022) %>% 
  
  # ## remove the timepoints that we do not have data for  
  # pink beads (round one)
  filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  filter(!(bead_color=="pink" & counted_when == "D2_aN")) %>%
  filter(!(bead_color=="pink" & counted_when == "D4_M")) %>%
  filter(!(bead_color=="pink" & counted_when == "D4_bN")) %>%
  # green beads (round two)
  filter(!(bead_color=="green" & counted_when == "D2_aN")) %>%
  filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  filter(!(bead_color=="green" & counted_when == "D4_M")) %>%
  # blue beads (round three)
  filter(!(bead_color=="blue" & counted_when == "D1_aN")) %>%
  filter(!(bead_color=="blue" & counted_when == "D2_bN")) %>%
  filter(!(bead_color=="blue" & counted_when == "D4_bN")) %>% 
  
  ## select columns to make them compatible with 2021 dataset
  select(colonyID, date,
         bead_color,
         age_2022, age_obs, age_class, 
         beads_collected, counted_when, cum_prop)
  
```

### Combine the 2021 and 2022 datasets

> Currently showing all the data from 2021 and 2022.

```{r combine_datasets}

## Combine the two data sets
tidydat <- rbind(old.summ, tidydat.2022) %>%
  mutate(counted_when = factor(counted_when,
                               levels = c("D1_bN", "D1_aN",
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
                            ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
                                   ifelse(age_obs >=5 & age_obs < 10, "5-9_yr",
                                                 "10+_yr")))) %>% 
  mutate(age_class = factor(age_class, 
                            levels = c("1-2_yr", "3-4_yr",
                                       "5-9_yr", "10+_yr")))

# # use only 2022 data
# tidydat <- tidydat.2022

which.colonies <- tidydat %>% select(colonyID, age_obs, age_class) %>% distinct()

tidydat <- tidydat %>% 
  mutate(colonyID = factor(colonyID, levels = which.colonies %>% arrange(age_obs) %>% pull(1) %>% unique()))

```



## Cumulative bead returned

```{r plot_boxplot}
# tidydat %>% 
#   
#   na.omit() %>% 
#   
#   ggplot(aes(x=counted_when, y=cum_prop, fill=age_class, color=age_class, group=age_class)) +
#   geom_jitter(size=2, alpha=0.4, width=0.2) +
#   # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
#   # geom_boxplot(alpha=0.3) +
#   geom_smooth(method = "lm", formula = y~log2(x+1), 
#               level=0.95,
#               alpha=0.4, size = 2,
#               aes(color=age_class)) +
#   # geom_smooth(method = "lm", formula = y~x, alpha=0.3, aes(color=age_class, lty=age_class)) +
#   
#   facet_wrap(~bead_color, nrow = 1) +
#   # coord_flip() +
#   # scale_y_continuous(n.breaks = 3,
#   #                    labels = c("","0.5", "1")) +
#   
#   theme_Publication(base_size = 15) +
#   # rotate x-axis lables
#   theme(axis.text.x=element_text(angle=90,hjust=1))+
#   # scale_fill_discrete() +
#   # scale_color_discrete()
# 
#   # scale_fill_viridis_d(option="cividis") +
#   # scale_color_viridis_d(option = "cividis")
#   
#   scale_fill_manual(values = age_colors) +
#   scale_color_manual(values = age_colors) +
#   
#   labs(x = "Day-time of observation",
#        y= "cumulative proportion \nof beads returned",
#        subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
#        caption = "Note 
#        there were flash floods on Day 2 evening
#        (Tuesday; we missed Wednesday's field work)
#        of the second experimental run (green beads),
#        which was three days prior to the start of the third round (blue beads).")


## All colonies - 1

tidydat %>% 
  
  na.omit() %>% 
  
  # ## remove the timepoints that we do not have data for  
  # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  
  ggplot(aes(x=counted_when, y=cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.4, width = 0.2) +
  # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
  # geom_boxplot(alpha=0.3, aes(group=age_class)) +
  # geom_violin(alpha=0.3) +
  # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
  geom_smooth(method = "lm", formula = y~log2(x+1), 
              alpha=0.4, size = 2,
              aes(color=age_class)) +
  
  # facet_grid(~age_class) +
  # coord_flip() +
  scale_y_continuous(n.breaks = 3,
                     labels = c("","0.5", "1")) +
  
  theme_Publication(base_size = 15) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
       caption = "Note
       data shown for all three rounds of bead experiment")


## All colonies - 2

tidydat[!is.infinite(rowSums(tidydat %>% select(cum_prop))),] %>% 
  
  # ## remove the timepoints that we do not have data for  
  # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  
  ggplot(aes(x=counted_when, y=cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.5, width = 0.2) +
  geom_line(aes(group=bead_color), size=2, alpha=0.7) +
  # geom_boxplot(alpha=0.6, aes(group=counted_when)) +
  # geom_violin(alpha=0.3) +
  # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
  
  # geom_smooth(method = "lm", formula = y~log2(x+1), 
  #             alpha=0.5, size = 2, 
  #             aes(color=age_class)) +
  
  facet_wrap(colonyID~bead_color, nrow=6) +
  # coord_flip() +
  # scale_y_continuous(n.breaks = 3,labels = c("","0.5", "1")) +
  
  # theme_Publication(base_size = 10) +
  theme_bw(base_size = 8) +
  # rotate x-axis lables
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank())+
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "",
       caption = "not normalized for foraging activity")
  
```


### Potential trash accumulated in colony

> The daily trash accumulated in the colony is proportional to the number of incoming foragers on a given day.


## Normalized rate of trash removal 

> cumulative number of beads returned / possible number of incoming foragers on Day 1

I am using the total number of foragers going in and out of the colony, for 30 secs, on Day 1 as a proxy for the possible number of incoming foragers.

```{r load_2022_foraging}
# set path to your working directory (change, if necessary)
path_to_repo2 = "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/01_field_season_2022/04_field_work/02_rprojects/01_checking_data"

# set path to the behavioral data within your repository 
path_to_data2 = "/01_data/"

## Specify the name of the master file as saved on your computer
filename2 <- "arizona_2022_master_list_Aug29_1440h.csv"

column.names <- c("date",
                  "was_it_raining",
                  "observer",
                  "colony_ID",
                  "first_obs",
                  "second_obs",
                  "time",
                  "in_1",
                  "out_1",
                  "in_2",
                  "out_2",
                  "in_3",
                  "out_3",
                  "in_4",
                  "out_4",
                  "in_5",
                  "out_5",
                  "in_6",
                  "out_6",
                  "in_7",
                  "out_7",
                  "in_8",
                  "out_8"
                  )

for.dat <- 
  read.table(paste0(path_to_repo2, path_to_data2, filename2),
                  sep = ",",
                  header = T,
                  na.strings = c(""," "),
                  stringsAsFactors = F,
                  skip = 1) %>% 
  as_tibble() %>% 
  mutate(date = as.POSIXct(date, format="%m/%d/%Y", tz="MST")) %>% 
  select(1:length(column.names))

colnames(for.dat) <- column.names

tidy.for.2022 <-
  for.dat %>% 
  # remove columns containing first and second obs
  select(-first_obs, -second_obs, -was_it_raining) %>% 
  
  
  group_by(date,observer,colony_ID,time) %>% 
  mutate(tot_in = sum(in_1,in_2,in_3,in_4,in_5,in_6,in_7,in_8, na.rm = T),
         tot_out = sum(out_1,out_2,out_3,out_4,out_5,out_6,out_7,out_8, na.rm=T)) %>% 
  mutate(foraging_30s = tot_in+tot_out) %>% 
  ungroup() %>% 
  
  mutate(colony_ID = as.factor(colony_ID),
         observer = as.factor(observer)) %>% 
  
  # na.omit() %>% 
  
  pivot_longer(.,
               cols = in_1:out_8,
               names_to = "inout",
               values_to = "n_ants") %>% 
  
  group_by(inout) %>% 
  mutate(inout2 = str_split_fixed(inout, "_", 2)[1]) %>%
  ungroup() %>% 
  
  select(date:time, inout, 
         inout2,
         n_ants, tot_in, tot_out, everything())

all.dates <- tidy.for.2022 %>% pull(date) %>% na.omit() %>% unique()
date.max <- tidy.for.2022 %>% pull(date) %>% na.omit() %>% max()
```


The task is to obtain the foraging (in and out) data for each of the bead colonies, for the duration of the different experimental rounds.

> pink: 8/16/2022; green: 8/22/22; blue: 8/26/22

```{r clean_foraging_data}
which.dates <- c(seq(as.Date("8/16/22", format="%m/%d/%y"), 
                    as.Date("8/19/22", format="%m/%d/%y"), by=1),
                 seq(as.Date("8/22/22", format="%m/%d/%y"), 
                     as.Date("8/25/22", format="%m/%d/%y"), by=1),
                 seq(as.Date("8/26/22", format="%m/%d/%y"), 
                     as.Date("8/29/22", format="%m/%d/%y"), by=1))
bead.dates <- data.frame(date = which.dates,
                         obs_day = factor(rep(c("D1", "D2", "D3", "D4"), 3),
                                          levels = c("D1", "D2", "D3", "D4")),
                         bead_color = factor(rep(c("pink", "green", "blue"), each=4),
                                             levels = c("pink", "green", "blue"))
                                             )
bead.colonies <- tidydat %>% pull(colonyID) %>% unique()
which.day.obs <- bead.dates %>% mutate(date_obs = paste0(date, obs_day)) %>% pull(date_obs)

tidy.for <- tidy.for.2022 %>% 
  
  mutate(date = as.character(date)) %>% 
  filter(date %in% unique(as.character(bead.dates$date))) %>% 
  mutate(date = as.Date(date)) %>%
  
  filter(colony_ID %in% bead.colonies) %>% 
  
  left_join(bead.dates, by=c("date")) %>% 

  select(date, bead_color, colony_ID, obs_day, time, tot_in, tot_out, foraging_30s) %>% 
  distinct() %>% 
  arrange(bead_color, colony_ID, date) %>% 
  ungroup()

tidy.for <-
  tidy.for %>%
  # calculate maximum foraging for each colony 
  group_by(colony_ID) %>% 
    summarise(max_foraging_30s = max(foraging_30s)) %>% 
    ungroup() %>% 
    right_join(tidy.for) %>% 
    mutate(prop_foraging_30s = foraging_30s/max_foraging_30s) %>% 
    mutate_if(is.numeric, round, 2)
    
  

s.tidydat.2022 <-
  tidydat.2022 %>% 
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% 
  full_join(tidy.for %>% select(-time) %>% select(colonyID=colony_ID, everything()) %>% filter(obs_day == "D1"), 
            by = c("date","bead_color","colonyID")) %>% 
  na.omit() %>% 
  mutate(s_cum_prop = round(cum_prop/foraging_30s, 2)) %>% 
  mutate(s2_cum_prop = round(cum_prop/prop_foraging_30s, 2)) %>% 
  na.omit() %>% 
  # filter(s_cum_prop <= 1) %>% 
  select(colonyID:bead_color,
         age_2022:age_class, 
         beads_collected,
         foraging_30s,
         counted_when, cum_prop, 
         s_cum_prop, s2_cum_prop)


## Load the foraging data for 2021 dataset
filename3 <- "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data/data/foraging_data.csv"

tidy.for.2021 <-
  read.csv(filename3) %>% as_tibble() %>%
    mutate(date = as.Date(date, format="%m/%d/%y")) %>%
    mutate(colonyID = as.factor(colonyID)) 
  
tidy.for.2021 <- 
  tidy.for.2021 %>% 
  ## identify max foragers for each colony
  group_by(colonyID) %>% 
  summarise(max_foraging_30s = max(foraging_30s)) %>% 
  ungroup() %>% 
  right_join(tidy.for.2021) %>% 
  mutate(prop_foraging_30s = foraging_30s/max_foraging_30s) %>% 
  mutate_if(is.numeric, round, 2)

s.tidydat.2021 <- 
  old.summ %>%
  mutate(colonyID = as.factor(colonyID)) %>% 
  mutate(obs_day = parse_number(as.character(counted_when))) %>% 
  filter(obs_day==1) %>% 
  left_join(tidy.for.2021, by=c("colonyID", "date")) %>% 
  select(colonyID:bead_color, obs_day, foraging_30s, prop_foraging_30s) %>% 
  
  mutate(colonyID = as.double(colonyID)) %>%
  right_join(old.summ, by=c("colonyID", "date", "bead_color")) %>% 
  
  group_by(colonyID,bead_color) %>%  
  mutate(foraging_30s = foraging_30s[1]) %>% 
  mutate(prop_foraging_30s = prop_foraging_30s[1]) %>% 
  mutate(obs_day = paste0("D",obs_day[1])) %>% 
  ungroup() %>% 
  
  mutate(s_cum_prop = round(cum_prop/foraging_30s, 2)) %>%
  mutate(s2_cum_prop = round(cum_prop/prop_foraging_30s, 2)) %>%
  
  na.omit() %>% 
  # filter(s_cum_prop <= 1) %>%
  
  select(colonyID:bead_color,
         age_2022:age_class, 
         beads_collected,
         foraging_30s,
         counted_when, cum_prop, 
         s_cum_prop, s2_cum_prop)


## combine the two files
s.tidydat <- 
  rbind(s.tidydat.2021, s.tidydat.2022) %>% 
    mutate(counted_when = factor(counted_when,
                               levels = c("D1_bN", "D1_aN",
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
                            ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
                                   ifelse(age_obs >=5 & age_obs < 10, "5-9_yr",
                                                 "10+_yr")))) %>% 
  mutate(age_class = factor(age_class, 
                            levels = c("1-2_yr", "3-4_yr",
                                       "5-9_yr", "10+_yr")))

## order the colony IDs according to age
s.tidydat <- s.tidydat %>% 
  mutate(colonyID = factor(colonyID, levels = which.colonies %>% arrange(age_obs) %>% pull(1) %>% unique()))
  

```


### Compare foraging across age-class

```{r plot_foraging_ageclass}
## Plot 1
s.tidydat %>% 
  select(colonyID, bead_color, age_obs, age_class, foraging_30s) %>% 
  distinct() %>%
  
  # group_by(age_obs) %>%
  # summarize(n_unique = length(unique(colonyID)))
  
  ggplot(aes(x=as.factor(age_obs), y=foraging_30s)) +
  
  geom_boxplot(alpha=0.7, outlier.color = NA) +
    
    geom_jitter(width = 0.1, size=3, alpha=0.6, 
                aes(fill=as.factor(colonyID)), 
                shape=21) +
  
    
    geom_smooth(method = "gam", aes(group=1)) +
    
    # stat_summary(fun = mean, geom = "point", col="black",
    #              size=10, alpha=0.8,
    #              position=position_nudge(x = 0.3, y = 0)) +
    # stat_summary(fun.data = mean_cl_boot,
    #              size=2, alpha = 0.8,
    #              geom = "errorbar", width = 0.2, fun.args = list(conf.int = 0.95),
    #              position=position_nudge(x = 0.3, y = 0)) +

    theme_Publication(base_size = 13) +

    labs(title = "Colony foraging across different age-groups",
       subtitle = "observations made on Day1 of bead expmts.",
       y = "activity \n(mean ± 95% CI)",
       x = "age-groups") +
  
    # scale_color_discrete() +
    # scale_fill_discrete() +

    # theme(legend.position="none") +

    theme(plot.subtitle=element_text(hjust=0.5, size=10),
          legend.title = element_blank()) +
    theme(legend.position="bottom")


## Plot 2
tidy.for %>% 
  select(date, colonyID=colony_ID, foraging_30s) %>% 
  right_join(which.colonies, by="colonyID") %>% 
  
  ggplot(aes(x=as.factor(age_obs), y=foraging_30s)) +
  
  geom_boxplot(alpha=0.7) +
    
    geom_jitter(width = 0.1, size=2, alpha=0.6, 
                aes(shape=as.factor(age_class), col=as.factor(age_class))) +
  
    
    stat_summary(fun.data = mean_cl_boot,
                 col="#F26C60",
                 size=2, alpha = 0.8,
                 geom = "errorbar", width = 0.2, fun.args = list(conf.int = 0.95),
                 position=position_nudge(x = 0.4, y = 0)) +
  stat_summary(fun = mean, geom = "line", 
               aes(group=1),
               col="#F26C60",
                 size=2, alpha=0.9,
                 position=position_nudge(x = 0.4, y = 0)) +
  # geom_smooth(method = "gam", aes(group=1)) +
  geom_smooth(method = "gam", aes(group=1), formula = y ~ splines::ns(x, 3)) +

    theme_Publication(base_size = 13) +

    labs(title = "Colony foraging across different age-groups",
       subtitle = "all observations made on bead colonies",
       y = "activity \n(mean ± 95% CI)",
       x = "age-groups") +
  
    # scale_color_discrete() +
    # scale_fill_discrete() +

    # theme(legend.position="none") +
    theme(plot.subtitle=element_text(hjust=0.5, size=10),
          legend.title = element_blank()) +
    theme(legend.position="bottom") +
  
    ## add text labels
    ggrepel::geom_label_repel(data = . %>% 
                    mutate(label = ifelse(foraging_30s > 100,
                                          colonyID, "")) %>% 
                    group_by(colonyID) %>% 
                    mutate_at(vars(label), funs(replace(., duplicated(.), ""))),
                    force = 11,
                    aes(label = label), 
                    box.padding = 0.3,
                    show.legend = FALSE)


## Plot 3
tidy.for.2022 %>% 
  mutate(colony_ID = as.numeric(as.character(colony_ID))) %>% 
  right_join(colony.dat %>% select(colony_ID = colonyID, age_2022)) %>% 
  na.omit() %>% 
  mutate(age_obs = age_2022) %>% 
  mutate(age_class = ifelse(age_obs <= 2, "1-2_yr",
                            ifelse(age_obs > 2 & age_obs <= 4, "3-4_yr",
                                   ifelse(age_obs >=5 & age_obs < 10, "5-9_yr",
                                                 "10+_yr")))) %>% 
  mutate(age_class = factor(age_class, 
                            levels = c("1-2_yr", "3-4_yr",
                                       "5-9_yr", "10+_yr"))) %>% 
  
  ## test labels
  # mutate(label = ifelse(foraging_30s > 200,
  #                                         colony_ID, "")) %>% 
  #                   group_by(colony_ID) %>%
  #                   mutate_at(vars(label), funs(replace(., duplicated(.), "")))
  
  # mutate(foraging_30s = log2(foraging_30s+1)) %>% 
  
  ggplot(aes(x=as.factor(age_obs), y=foraging_30s)) +
  geom_boxplot(alpha=0.4, fill="#C5E4BA") +
  geom_jitter(width = 0.2, size=1, alpha=0.2, 
                aes(shape=as.factor(age_class), col=as.factor(age_class))) +
  stat_summary(fun.data = mean_cl_boot,
                 col="#F26C60",
                 size=2, alpha = 0.8,
                 geom = "errorbar", width = 0.2, fun.args = list(conf.int = 0.95),
                 position=position_nudge(x = 0.4, y = 0)) +
  stat_summary(fun = mean, geom = "line", 
               aes(group=1),
               col="#F26C60",
                 size=2, alpha=0.9,
                 position=position_nudge(x = 0.4, y = 0)) +
  geom_smooth(method = "gam", aes(group=1)) +
    theme_Publication(base_size = 13) +
    labs(title = "Colony foraging across different age-groups",
       subtitle = "all observations of colony foraging in 2022",
       y = "activity \n(mean ± 95% CI)",
       x = "age-groups",
       caption = "red line indicates mean ± 95% CI; \nblue line shows a GAM fit; \nlabels indicate colonies with activity more than 150 ants/30 secs") +
    # ylim(c(0,200)) +
    theme(plot.subtitle=element_text(hjust=0.5, size=10),
          plot.caption = element_text(hjust=1, size=10),
          legend.title = element_blank()) +
    theme(legend.position="bottom") +
  
    ## add text labels
    ggrepel::geom_label_repel(data = . %>% 
                    mutate(label = ifelse(foraging_30s > 150,
                                          colony_ID, "")) %>% 
                    group_by(colony_ID) %>% 
                    mutate_at(vars(label), funs(replace(., duplicated(.), ""))),
                    force = 1,
                    size=2.5,
                    aes(label = label), 
                    box.padding = 0.3,
                    show.legend = FALSE)

```


### Plot normalized bead returns

```{r plot_normalized_bead_returns}
# s.tidydat[!is.infinite(rowSums(s.tidydat %>% select(s_cum_prop, s2_cum_prop))),] %>% 
#   
#   ggplot(aes(x=counted_when, y=s_cum_prop, fill=age_class, color=age_class, group=age_class)) +
#   geom_jitter(size=2, alpha=0.4, width=0.2) +
#   # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
#   # geom_boxplot(alpha=0.3) +
#   geom_smooth(method = "lm", formula = y~log2(x+1), 
#               level=0.95,
#               alpha=0.4, size = 2,
#               aes(color=age_class)) +
#   # geom_smooth(method = "lm", formula = y~x, alpha=0.3, aes(color=age_class, lty=age_class)) +
#   
#   facet_wrap(~bead_color, nrow = 1) +
#   # coord_flip() +
#   # scale_y_continuous(n.breaks = 3, labels = c("","0.5", "1")) +
#   
#   theme_Publication(base_size = 15) +
#   # rotate x-axis lables
#   theme(axis.text.x=element_text(angle=90,hjust=1))+
#   # scale_fill_discrete() +
#   # scale_color_discrete()
# 
#   # scale_fill_viridis_d(option="cividis") +
#   # scale_color_viridis_d(option = "cividis")
#   
#   scale_fill_manual(values = age_colors) +
#   scale_color_manual(values = age_colors) +
#   
#   labs(x = "Day-time of observation",
#        y= "cumulative proportion \nof beads returned",
#        subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
#        caption = "Note 
#        there were flash floods on Day 2 evening
#        (Tuesday; we missed Wednesday's field work)
#        of the second experimental run (green beads),
#        which was three days prior to the start of the third round (blue beads).")


## All colonies - 1

s.tidydat[!is.infinite(rowSums(s.tidydat %>% select(s_cum_prop, s2_cum_prop))),] %>% 
  
  ggplot(aes(x=counted_when, y=s_cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.4, width = 0.2) +
  # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
  # geom_boxplot(alpha=0.3, aes(group=age_class)) +
  # geom_violin(alpha=0.3) +
  # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
  geom_smooth(method = "lm", formula = y~log2(x+1), 
              alpha=0.4, size = 2,
              aes(color=age_class)) +
  
  theme_Publication(base_size = 15) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
       caption = "normalized by foraging activity on D1 of bead experiments")


## All colonies - 2

s.tidydat[!is.infinite(rowSums(s.tidydat %>% select(s_cum_prop, s2_cum_prop))),] %>% 
  
  # ## remove the timepoints that we do not have data for  
  # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  
  ggplot(aes(x=counted_when, y=s_cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.5, width = 0.2) +
  geom_line(aes(group=bead_color), size=2, alpha=0.7) +
  # geom_boxplot(alpha=0.6, aes(group=counted_when)) +
  # geom_violin(alpha=0.3) +
  # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
  
  # geom_smooth(method = "lm", formula = y~log2(x+1), 
  #             alpha=0.5, size = 2, 
  #             aes(color=age_class)) +
  
  facet_wrap(colonyID~bead_color, nrow=6) +
  # coord_flip() +
  # scale_y_continuous(n.breaks = 3,labels = c("","0.5", "1")) +
  
  # theme_Publication(base_size = 10) +
  theme_bw(base_size = 8) +
  # rotate x-axis lables
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank())+
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "",
       caption = "normalized by foraging activity on D1 of bead experiments")

```


```{r plot_with_age}
# > I am wondering how would it look if I plot the curves for each age.

# s.tidydat %>% 
#   
#   # ## remove the timepoints that we do not have data for  
#   # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
#   # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
#   
#   ggplot(aes(x=counted_when, y=s_cum_prop, 
#              fill=as.factor(age_obs), color=as.factor(age_obs), group=as.factor(age_obs))) +
#   geom_jitter(size=2, alpha=0.4, width = 0.2) +
#   # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
#   # geom_boxplot(alpha=0.3, aes(group=age_class)) +
#   # geom_violin(alpha=0.3) +
#   # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
#   geom_smooth(method = "lm", formula = y~log2(x+1), 
#               alpha=0.4, size = 2,
#               aes(color=as.factor(age_obs))) +
#   
#   facet_grid(~age_class) +
#   # coord_flip() +
#   # scale_y_continuous(n.breaks = 3, labels = c("","0.5", "1")) +
#   
#   theme_Publication(base_size = 15) +
#   # rotate x-axis lables
#   theme(axis.text.x=element_text(angle=90,hjust=1))+
#   
#   scale_fill_viridis_d() +
#   scale_color_viridis_d() +
#   
#   labs(x = "Day-time of observation",
#        y= "cumulative proportion \nof beads returned",
#        subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
#        caption = "Note
#        data shown for all three rounds of bead experiment")
```


```{r plot_with_age_2}
# > What about colony variation withih each age-class

# s.tidydat %>% 
#   
#   # ## remove the timepoints that we do not have data for  
#   # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
#   # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
#   
#   ggplot(aes(x=counted_when, y=s_cum_prop, 
#              fill=as.factor(colonyID), color=as.factor(colonyID), group=as.factor(colonyID))) +
#   geom_jitter(size=2, alpha=0.4, width = 0.2) +
#   # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
#   # geom_boxplot(alpha=0.3, aes(group=age_class)) +
#   # geom_violin(alpha=0.3) +
#   # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
#   geom_smooth(method = "lm", formula = y~log2(x+1), 
#               alpha=0.4, size = 2,
#               aes(color=as.factor(colonyID))) +
#   
#   facet_grid(~age_class) +
#   # coord_flip() +
#   # scale_y_continuous(n.breaks = 3, labels = c("","0.5", "1")) +
#   
#   theme_Publication(base_size = 15) +
#   # rotate x-axis lables
#   theme(axis.text.x=element_text(angle=90,hjust=1))+
#   
#   scale_fill_viridis_d() +
#   scale_color_viridis_d() +
#   
#   labs(x = "Day-time of observation",
#        y= "cumulative proportion \nof beads returned",
#        subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
#        caption = "Note
#        data shown for all three rounds of bead experiment")
```


### Plot normalized bead returns v2

> Standardizing by proportion of foragers outside on Day 1 of bead experiments

The idea is to divide the cumulative proportion of beads returned by the proportion of foragers (among the max foragers that were observed for a given colony) out on Day 1 of the bead experiments 

```{r standardizing_v2}
# s.tidydat[!is.infinite(rowSums(s.tidydat %>% select(s_cum_prop, s2_cum_prop))),] %>% 
#   
#   ggplot(aes(x=counted_when, y=s2_cum_prop, fill=age_class, color=age_class, group=age_class)) +
#   geom_jitter(size=2, alpha=0.4, width=0.2) +
#   # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
#   # geom_boxplot(alpha=0.3) +
#   geom_smooth(method = "lm", formula = y~log2(x+1), 
#               level=0.95,
#               alpha=0.4, size = 2,
#               aes(color=age_class)) +
#   # geom_smooth(method = "lm", formula = y~x, alpha=0.3, aes(color=age_class, lty=age_class)) +
#   
#   facet_wrap(~bead_color, nrow = 1) +
#   # coord_flip() +
#   # scale_y_continuous(n.breaks = 3, labels = c("","0.5", "1")) +
#   
#   theme_Publication(base_size = 15) +
#   # rotate x-axis lables
#   theme(axis.text.x=element_text(angle=90,hjust=1))+
#   # scale_fill_discrete() +
#   # scale_color_discrete()
# 
#   # scale_fill_viridis_d(option="cividis") +
#   # scale_color_viridis_d(option = "cividis")
#   
#   scale_fill_manual(values = age_colors) +
#   scale_color_manual(values = age_colors) +
#   
#   labs(x = "Day-time of observation",
#        y= "cumulative proportion \nof beads returned",
#        subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
#        caption = "")


## All colonies - 1

s.tidydat[!is.infinite(rowSums(s.tidydat %>% select(s_cum_prop, s2_cum_prop))),] %>% 
  
  # ## remove the timepoints that we do not have data for  
  # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  
  ggplot(aes(x=counted_when, y=s2_cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.4, width = 0.2) +
  # geom_line(aes(group=colonyID), size=0.5, alpha=0.7) +
  # geom_boxplot(alpha=0.3, aes(group=age_class)) +
  # geom_violin(alpha=0.3) +
  # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
  geom_smooth(method = "lm", formula = y~log2(x+1), 
              alpha=0.4, size = 2,
              aes(color=age_class)) +
  
  # facet_grid(~age_class) +
  # coord_flip() +
  # scale_y_continuous(n.breaks = 3, labels = c("","0.5", "1")) +
  
  theme_Publication(base_size = 15) +
  # rotate x-axis lables
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "Line indicates y ~ log2(x+1) + 95% CI",
       caption = "")


## All colonies - 2
s.tidydat[!is.infinite(rowSums(s.tidydat %>% select(s_cum_prop, s2_cum_prop))),] %>% 
  # ## remove the timepoints that we do not have data for  
  # filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  # filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  ggplot(aes(x=counted_when, y=s2_cum_prop, fill=age_class, color=age_class, group=age_class)) +
  geom_jitter(size=2, alpha=0.2, width = 0.2) +
  geom_line(aes(group=bead_color), size=1, alpha=0.7) +
  # geom_boxplot(alpha=0.6, aes(group=counted_when)) +
  # geom_violin(alpha=0.3) +
  # geom_smooth(method = "gam", formula=y~s(x, k=5)) +
  # geom_smooth(method = "lm", formula = y~log2(x+1), 
  #             alpha=0.5, size = 2, 
  #             aes(color=age_class)) +
  facet_wrap(colonyID~bead_color, nrow=6) +
  # coord_flip() +
  # scale_y_continuous(n.breaks = 3,labels = c("","0.5", "1")) +
  
  # theme_Publication(base_size = 8) +
  theme_bw(base_size = 8) +
  # remove x-axis labels and ticks
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank())+
  
  scale_fill_manual(values = age_colors) +
  scale_color_manual(values = age_colors) +
  
  labs(x = "Day-time of observation",
       y= "cumulative proportion \nof beads returned",
       subtitle = "",
       caption = "normalized by proportion of foragers out on D1 of bead experiments")

```


___

## Prelim Analyses I

```{r prelim_analyses_v1}

s.tidydat.mod <- s.tidydat %>% mutate(log_s_cum_prop = log2(s_cum_prop+1))
s.tidydat.mod <- s.tidydat.mod[!is.infinite(rowSums(s.tidydat.mod %>% select(s_cum_prop, s2_cum_prop))),]


mod.full <- lme4::lmer(log_s_cum_prop ~ age_class + counted_when + (1|colonyID) + (1|bead_color) ,
                 data=s.tidydat.mod, REML=F)
mod.red.1 <- lme4::lmer(log_s_cum_prop ~ counted_when + (1|colonyID) + (1|bead_color) ,
                 data=s.tidydat.mod, REML=F)
anova(mod.full, mod.red.1, test="Chisq")


## Let's look at the summary of the model
lmerTest::ranova(mod.full)

## Calculate the p-values for the fixed effects
## Source: https://mspeekenbrink.github.io/sdam-r-companion/linear-mixed-effects-models.html
afex::mixed(log_s_cum_prop ~ age_class + counted_when + (1|colonyID) + (1|bead_color) ,
                 data=s.tidydat.mod, check_contrasts = T, test_intercept = F, method="KR")

```


## Prelim Analyses II

```{r prelim_analyses_v2}

s.tidydat.mod <- s.tidydat %>% mutate(log_s2_cum_prop = log2(s2_cum_prop+1))
s.tidydat.mod <- s.tidydat.mod[!is.infinite(rowSums(s.tidydat.mod %>% select(s_cum_prop, s2_cum_prop))),]


mod.full <- lme4::lmer(log_s2_cum_prop ~ age_class + counted_when + (1|colonyID) + (1|bead_color) ,
                 data=s.tidydat.mod, REML=F)
mod.red.1 <- lme4::lmer(log_s2_cum_prop ~ counted_when + (1|colonyID) + (1|bead_color) ,
                 data=s.tidydat.mod, REML=F)
anova(mod.full, mod.red.1, test="Chisq")


## Let's look at the summary of the model
lmerTest::ranova(mod.full)

## Calculate the p-values for the fixed effects
## Source: https://mspeekenbrink.github.io/sdam-r-companion/linear-mixed-effects-models.html
afex::mixed(log_s2_cum_prop ~ age_class + counted_when + (1|colonyID) + (1|bead_color) ,
                 data=s.tidydat.mod, check_contrasts = T, test_intercept = F, method="KR")
```

## Other analyses

> Run GAM instead of lme4::lmer()

[read more here](https://fromthebottomoftheheap.net/2021/02/02/random-effects-in-gams/)

___

## Extras

> How many colonies, and which colonies, do we have for each age-class? 

```{r tallying_22Aug22_green_beads}
s.tidydat %>% 
  mutate(age_obs = factor(paste0(age_obs, " yr"), 
                          level=paste0(sort(unique(.$age_obs)), " yr"))) %>%
  group_by(age_obs) %>% 
  summarize(n_colonies = length(unique(colonyID)),
            which_colonies = paste(as.character(unique(colonyID)), collapse = ", ")) %>% 
  
  kableExtra::kable(caption = "Colonies used in bead experiments", align = "c") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
  

```







