
---
title: "Daily data check"
author:
- name: foraging count data for 2022 field work
  affiliation: In case of questions, please contact [Biplabendu Das](https://biplabendu.github.io/homepage/)
date: Last updated on `r format(Sys.time(), "%d %B, %Y")`
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    keep_md: no
    code_folding: hide
    includes:
      in_header: analytics.html
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

set.seed(420)
rm(list = ls())

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

```


___

## Housekeeping

```{r housekeeping, include=FALSE}

# Load the libraries
pacman::p_load(tidyverse, conflicted, ggplot2, viridis, conflicted, WaveletComp)
pacman::p_load(patchwork)
pacman::p_load(lubridate, kableExtra)

# set conflict preference
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("union", "dplyr")
conflict_prefer("lag", "dplyr")

# set path to your working directory (change, if necessary)
path_to_repo = "/Users/biplabendudas/OneDrive - University of Central Florida/UCF de bekker/PostDoc/Labs/Gordon/04_research/2022/01_field_season_2022/04_field_work/02_rprojects/01_checking_data"

# set path to the behavioral data within your repository 
path_to_data = "/01_data/"


# set path to the results
path_to_results = "/Users/biplabendudas/Documents/GitHub/homepage/images/2022_deborah/"

# load functions
# customized theme for publication quality figures
source(paste0(path_to_repo,"/03_functions/theme_publication.R"))
source(paste0(path_to_repo,"/03_functions/plot_zscores.R"))

```


```{r user_input, include=F}

## This is where you want to specify your parameters of choice.

## Specify the name of the master file as saved on your computer
filename <- "arizona_2022_master_list_Aug29_1440h.csv"

column.names <- c("date",
                  "was_it_raining",
                  "observer",
                  "colony_ID",
                  "first_obs",
                  "second_obs",
                  "time",
                  "in_1",
                  "out_1",
                  "in_2",
                  "out_2",
                  "in_3",
                  "out_3",
                  "in_4",
                  "out_4",
                  "in_5",
                  "out_5",
                  "in_6",
                  "out_6",
                  "in_7",
                  "out_7",
                  "in_8",
                  "out_8"
                  )

## Read the age data for the colonies
colony.dat <- 
  read.csv(paste0(path_to_repo,
                              "/../../03_datasheets/from_Deborah/",
                              "foraging_counts_08_13_22.csv"),
                       header = T, stringsAsFactors = F) %>% 
  as_tibble() %>% 
  select(quad:age_2022) %>% 
  mutate(colony_ID = as.factor(colony_ID)) %>% 
  mutate(age_class = ifelse(age_2022 <= 2, "young",
                           ifelse(age_2022 > 2 & age_2022 <= 4, "teen",
                                  ifelse(age_2022 > 4 & age_2022 <= 9, "young_adult",
                                         ifelse(age_2022 > 9 & age_2022 <= 17, "adult",
                                                "old"))))) %>% 
  mutate(age_class = factor(age_class, 
                            levels = c("young","teen","young_adult","adult","old")))


## read the bead data for 2022
bead.dat <- read.csv("/Users/biplabendudas/OneDrive - University of Central Florida/UCF de bekker/PostDoc/Labs/Gordon/00_application_package/final_round/data/bead_data_2022/all_bead_data_2022/2022_bead_data_master_replicate_one_two_three_26Aug22.csv", header = T, stringsAsFactors = F) %>% as_tibble()
# which are the bead colonies?
bead.colonies <- bead.dat %>% pull(colonyID) %>% unique() %>% na.omit() %>% as.character()

```

___

## 1. Read data and tidy

```{r tidy_data}

dat <- 
  read.table(paste0(path_to_repo, path_to_data, filename),
                  sep = ",",
                  header = T,
                  na.strings = c(""," "),
                  stringsAsFactors = F,
                  skip = 1) %>% 
  as_tibble() %>% 
  mutate(date = as.POSIXct(date, format="%m/%d/%Y", tz="MST")) %>% 
  select(1:length(column.names))

colnames(dat) <- column.names

tidy.dat <-
  dat %>% 
  # remove columns containing first and second obs
  select(-first_obs, -second_obs, -was_it_raining) %>% 
  
  
  group_by(date,observer,colony_ID) %>% 
  mutate(tot_in = sum(in_1,in_2,in_3,in_4,in_5,in_6,in_7,in_8, na.rm = T),
         tot_out = sum(out_1,out_2,out_3,out_4,out_5,out_6,out_7,out_8, na.rm=T)) %>% 
  ungroup() %>% 
  
  mutate(colony_ID = as.factor(colony_ID),
         observer = as.factor(observer)) %>% 
  
  # na.omit() %>% 
  
  pivot_longer(.,
               cols = in_1:out_8,
               names_to = "inout",
               values_to = "n_ants") %>% 
  
  group_by(inout) %>% 
  mutate(inout2 = str_split_fixed(inout, "_", 2)[1]) %>%
  ungroup() %>% 
  
  select(date:time, inout, 
         inout2,
         n_ants, tot_in, tot_out, everything())

all.dates <- tidy.dat %>% pull(date) %>% na.omit() %>% unique()
date.max <- tidy.dat %>% pull(date) %>% na.omit() %>% max()
```

Analyzing data from 8/13/2022 (Aug 13, 2022) until `r date.max`.

> Currently showing all data for: `r date.max`

## 2. Summarize data

```{r summary_table}
## Who did what?
tidy.dat %>% 
  filter(date == date.max) %>% 
  select(date:colony_ID) %>% 
  distinct() %>% 
  # add colony age data
  left_join(colony.dat, by="colony_ID") %>% 
  # summarize the number of colonies per age class
  group_by(observer, age_class, date) %>% 
  summarize(n_colonies = length(unique(colony_ID))) %>% 
  ungroup() %>% 
  # make data wide-formatted
  pivot_wider(names_from = age_class, values_from = n_colonies) %>% 
  replace_na(list(young=0, teen=0, young_adult=0, adult=0, old=0)) %>% 
  # rearrange columns
  select(observer,
         old, adult, young_adult, teen, young) %>% 
  
  # Style for html output
  kable(caption = paste("Table 1: Summary table for all foraging data collected on:",
                        date.max)) %>% 
  kable_styling(bootstrap_options = c("striped"), 
                full_width = F, position = "left", latex_options = "scaled_down")
```


___

## 2. Visualize as boxplots

```{r plot_boxplots, include=F}
observers <-
  tidy.dat %>% 
  filter(date == date.max) %>% 
  select(date:time, n_ants) %>% 
  na.omit() %>% 
  pull(observer) %>% unique() %>% na.omit() %>% as.character()

plot.x <- list()

for(i in 1:length(observers)) {
  
  colony.order <-
    tidy.dat %>% 
    filter(date == date.max) %>% 
    
    filter(observer == observers[[i]]) %>% 
    
    mutate(time = as.POSIXct(time, format="%H:%M")) %>% 
    arrange(date, time) %>% 
    na.omit() %>% 
    pull(colony_ID) %>% 
    unique() %>% 
    as.character()
  
  plot.x[[i]] <- 
    tidy.dat %>%
    filter(date == date.max) %>% 
      filter(observer == observers[[i]]) %>%
      mutate(colony_ID = factor(colony_ID, levels = colony.order)) %>%
      filter(!is.na(colony_ID)) %>% 
    ggplot(aes(x=colony_ID, y=n_ants)) +
      geom_boxplot(aes(fill=inout2), 
                   alpha=0.75, 
                   outlier.size = 3,
                   outlier.shape = 21,
                   outlier.color = "red",
                   position = position_dodge(width = 0.95)) +
      theme_Publication(base_size = 40) +
      scale_fill_Publication() +
      coord_flip() +
    labs(x=paste0(observers[[i]], "'s colony ID"))

  
}

png(paste0(path_to_results,
           "00_combined_images/",
           "all_boxplots.png"), 
      width = 40, height = 180, units = "cm", res = 120)
multi.plot(plot.x, rows = 5, cols=2)
trash <- dev.off()

```



```{r embed_boxplots_all, echo=F, out.width="75%", fig.cap='', fig.align='left'}

knitr::include_graphics(paste0(path_to_results,
           "00_combined_images/",
           "all_boxplots.png"))
  
```



## 3. Foraging with time

> In the following figure, the y-axis shows the ratio of out-going to in-going foragers, and the x-axis shows the colonies that were observed by each person. The colonies are arranged in the order in which they were observed.


```{r foraging_time, include=F}
plot.x <- list()

for(i in 1:length(observers)) {
  
  f.name <- paste0(observers[[i]],"_boxplots.png")
  
  colony.order <-
    tidy.dat %>% 
    filter(date == date.max) %>% 
    
    filter(observer == observers[[i]]) %>% 
    
    mutate(time = as.POSIXct(time, format="%H:%M")) %>% 
    arrange(date, time) %>% 
    na.omit() %>% 
    pull(colony_ID) %>% 
    unique() %>% 
    as.character()
    
  plot.x[[i]] <-
    tidy.dat %>% 
      filter(date == date.max) %>% 
      filter(observer == observers[[i]]) %>%
      mutate(colony_ID = factor(colony_ID, levels = colony.order)) %>% 
      select(date:time, tot_in, tot_out) %>% 
      distinct() %>% 
      na.omit() %>% 
      mutate(out_in_ratio = round(tot_out/tot_in, 2)) %>% 
    ## plot
    ggplot(aes(x=colony_ID, y=out_in_ratio)) +
      geom_hline(yintercept = 1, color="red", linetype=19, size=2, alpha=0.5) +
      geom_line(aes(group=1), size=3) +
      scale_y_continuous(breaks = c(0,1,2)) +
      theme_Publication(base_size = 60) +
      scale_fill_Publication() +
      # coord_flip() +
    labs(x=paste0(observers[[i]], "'s colonies; time -->"),
         y="out-to-in ratio") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
  # png(paste0(path_to_results,
  #          f.name), 
  #     width = 30, height = 45, units = "cm", res = 300)
  # print(plot.x[[i]])
  # trash <- dev.off()
  
}

png(paste0(path_to_results,
           "00_combined_images/",
           "io_ratio_with_time.png"), 
      width = 80, height = 180, units = "cm", res = 120)
multi.plot(plot.x, rows = 9 , cols=1)
trash <- dev.off()
```



```{r embed_foraging_time, echo=F, out.width="85%", fig.cap='', fig.align='left'}

knitr::include_graphics(paste0(path_to_results,
           "00_combined_images/",
           "io_ratio_with_time.png"))
  
```

___

## Check consistency of data collection

> Check if the colonies that were observed (obtained foraging counts) the previous day were also observed today.


```{r for_deborah_4}
n.day <- length(all.dates)

colonies.counting.today <- 
  dat %>% 
  filter(date==all.dates[n.day]) %>%
  select(observer, colony_ID) %>% 
  na.omit() %>%
  pull(colony_ID) %>% 
  unique() %>%
  as.character()


colonies.counting.previousday <- 
  dat %>% 
  filter(date==all.dates[n.day-1]) %>%
  select(observer, colony_ID) %>% 
  na.omit() %>%
  pull(colony_ID) %>% 
  unique() %>%
  as.character()

## Today v. previous day
writeLines(paste0("Colonies observed on ", all.dates[n.day], " but NOT on ", all.dates[n.day-1]))
dplyr::setdiff(colonies.counting.today, colonies.counting.previousday)
writeLines(paste0("Observed on ", all.dates[n.day-1], " but NOT on ", all.dates[n.day]))
dplyr::setdiff(colonies.counting.previousday, colonies.counting.today)
writeLines(paste0("Bead-colonies that do not have foraging count data for ", all.dates[n.day], " :"))
dplyr::setdiff(bead.colonies, colonies.counting.today)


```

___

## Extra stuff

```{r assign_colonies_22Aug22, eval=F, echo=F}
dat %>% 
  filter(date == date.max) %>% 
  select(observer, colony_ID) %>% 
  group_by(observer) %>% 
  summarise(num_colonies = n())

```


```{r timecourse_1, eval=F, echo=F}
previous.colonies <- c("833","859","863","1017",
                       "1068","1107","1242","1317","1368",
                       "757","664","509",
                       "1010", # also called "10"
                       "758")

dat %>% filter(colony_ID %in% previous.colonies) %>% 
  arrange(colony_ID, date) %>% 
  group_by(colony_ID) %>% 
  summarize(num_days_observed = n())

```


```{r for_deborah_5, eval=F, echo=F}
active.dat <- 
  tidy.dat %>% 
  na.omit() %>% 
  filter(date != date.max) %>% 
  group_by(colony_ID, date) %>% 
  summarise(n()) %>% 
  ungroup() %>% 
  group_by(colony_ID) %>% 
  summarize(num_days_counted_foraging = n()) %>% 
  arrange(num_days_counted_foraging)

writeLines("Are there colonies for which there is count data, but no such colony ID exists?")
all.colonies <- colony.dat %>% pull(colony_ID) %>% unique() %>% na.omit() %>% as.character()
active.dat %>% 
  pull(colony_ID) %>% 
  unique() %>% na.omit() %>% as.character() %>% 
  dplyr::setdiff(., all.colonies)

# active.dat %>% 
#   filter(!colony_ID %in% c("1312")) %>% 
#   left_join(colony.dat %>% select(1:2), by="colony_ID") %>% 
#   arrange(quad) %>% 
#   write.csv(.,
#             file = paste0(path_to_repo, "/04_results/",
#                           "colonies_for_which_foraging_data_is_available_until_27Aug22.csv"),
#             row.names = F)

```



```{r for_deborah, eval=F, echo=F}

## Summary table for Deborah
tidy.dat %>% 
  filter(date == date.max) %>% 
  select(date, colony_ID, observer) %>% 
  distinct() %>% 
  left_join(colony.dat, by="colony_ID") %>% 
  mutate(age_class = ifelse(age_2022 <= 2, "1-2_yr_old",
                                  ifelse(age_2022 > 2 & age_2022 <= 4, "3-4_yr_old",
                                         "older_than_5_yr"))) %>% 
  filter(age_class == "3-4_yr_old") %>% 
  pull(colony_ID)
  # # save this to a csv file
  # write.csv(.,
  #           file = paste0(path_to_repo, 
  #                         "/04_results/", 
  #                         "colony_ID_age_counted_on_08_14_2022.csv"),
  #           row.names = F)
  
  # group_by(age_class) %>%
  # summarise(n_colonies = n())
  
```


```{r bd_colonies, eval=F, echo=F}

bd.colonies <- c("1461", "1539", "1270",
                 "1340", "1417", "1543",
                 "1272", "1377", "1048",
                 "1157", "1556", "1324",
                 "1551", "1294", "1420",
                 "1262", "1245", "1366")

bd.dat <- 
  tidy.dat %>% 
  filter(colony_ID %in% bd.colonies) %>% 
  select(date:time) %>% 
  distinct() %>% 
  arrange(colony_ID) %>% 
  na.omit() %>% 
  left_join(colony.dat, by="colony_ID") %>% 
  mutate(age_class = ifelse(age_2022 <= 2, "1-2_yr_old",
                                  ifelse(age_2022 > 2 & age_2022 <= 4, "3-4_yr_old",
                                         "older_than_5_yr"))) 
  # bd.dat %>% 
  #   group_by(age_class) %>%
  #   summarise(n_colonies = length(unique(colony_ID)))


# Q. Which high priority colonies have I not looked at yet?
dplyr::setdiff(bd.colonies,
               unique(bd.dat$colony_ID))
# A: Colony 1366

# Q. Do we have some data for colony 1366?
dat %>% filter(colony_ID == "1366")
# A: Yes, MNW observed 1366 two times on 14 aug 2022; ants were somewhat active (S) around 7:53 AM, but not foraging at all (Z) at 10:18 AM.
```


```{r for_deborah_2, eval=F, echo=F}
active.colony.dat <- read.csv(paste0(path_to_repo, path_to_data, "active_colonies_8_15.csv"),
                              stringsAsFactors = F, header=T) %>% as_tibble()
active_8_15 <-
  active.colony.dat %>% 
  select(colony_ID, age, active_8_15) %>% 
  mutate(age = as.numeric(age)) %>% 
  filter(age > 5) %>% 
  filter(active_8_15 == "A") %>% 
    pull(colony_ID) %>% 
    unique() %>% 
    as.character()

counted_8_14 <-
  tidy.dat %>% 
    filter(date == all.dates[2]) %>% 
    pull(colony_ID) %>% 
    unique() %>% 
    as.character()

counted_8_15 <-
  tidy.dat %>% 
    filter(date == all.dates[3]) %>% 
    pull(colony_ID) %>% 
    unique() %>% 
    as.character()

# Colonies active but not counted on 8_14
dplyr::setdiff(active_8_15, counted_8_14)
dplyr::setdiff(active_8_15, counted_8_15)
   

# # check some colonies
# some.colonies <-  c("629","812","1107",
#                     "1144",
#                     "1195", "1281", "1329", "1403", "1568")

some.colonies <- c("1107")
tidy.dat %>%
  filter(date == all.dates[3]) %>%
  filter(colony_ID %in% some.colonies)

active.colony.dat %>%
  # filter(colony_ID %in% some.colonies)
  filter(colony_ID %in% some.colonies)
```


```{r for_deborah_3, eval=F, echo=F}
bd.colonies <- c("1461", "1539", "1270",
                 "1543", "1417", "1340",
                 "1272", "1377", "1048",
                 "1324", "1556", "1157",
                 "1551", "1294", "1420",
                 "1366", "1245", "1262",
                 ### new assigned
                 "1540", "1545")


tidy.dat %>% 
  filter(colony_ID %in% bd.colonies) %>% 
  select(colony_ID) %>% 
  distinct() %>% 
  left_join(colony.dat, by="colony_ID") %>% 
  mutate(age_class = ifelse(age_2022 <= 2, "1-2_yr_old",
                                  ifelse(age_2022 > 2 & age_2022 <= 4, "3-4_yr_old",
                                         "older_than_5_yr"))) %>%
  arrange(age_2022) 
```

___

## Supplementary

#### 1. Reproducibility

- All code used for this document is available on my [Github](https://github.com/biplabendu/homepage/blob/master/checking_data_swrs_2022.Rmd).
