---
title: "Who is attending IUSSI?"
author: Biplabendu Das
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    theme: yeti
    includes:
      in_header: analytics.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

```

***

> Data: note from Ben Oldroyd (President, IUSSI 2014)

"For the Cairns meeting, everyone who submitted an abstract and paid the registration (or had it paid by IUSSI) was allowed to speak or give a poster as per their stated preference. Everyone who submitted a symposium proposal got to chair/organize that symposium (in one or two cases we asked proposers to team up)."

> What is this document?   

The goal was to explore the data for the past participants of the internation IUSSI conferences (2014 and 2018) to get a sense of the demography of the representatives. Note, this data is for the final list of participants that attended the conferences and not a comprehensive list of all applicants (see note from Ben Oldroyd). Please note, this is still a work in progress, and the document will change with time.


```{r, include=F, echo=F}
# Let's load the data and get a glimpse.

rm(list = ls())
library(tidyverse)
library(countrycode)
dat.2014 <- read.csv("./data/2014_participants_list_final.csv", header = T, stringsAsFactors = F)

# Get a glimpse
dat.2014 %>% glimpse()

# Let's summarize the paricipants by country
dat.per.country.2014 <- 
  dat.2014 %>% group_by(country) %>% 
  summarise(participants_2014 = n()) %>% arrange(desc(participants_2014)) %>% 
  # country names are in different formats;change them to a common one (iso3c format)
  mutate(country.iso3c = countrycode(country,origin='country.name', destination='iso3c'))

# How many countries were present?
dat.per.country.2014 %>% nrow()

# Any unknown countries?
# dat.per.country.2014 %>% filter(is.na(country.iso3c)) # nope

# What's the median number of participants per country?
dat.per.country.2014 %>% pull(participants_2014) %>% median()

# Let's take a look at the top 15 and bottom 15 countries by participation
dat.per.country.2014 %>% head(15)
dat.per.country.2014 %>% tail(15)


```



```{r, inlcude=F, echo=F}
## Let's compare 2014 and 2018
dat.2018 <- read.csv("./data/2018_participants_list.csv", header = T, stringsAsFactors = F)

# # Get a glimpse
# dat.2018 %>% glimpse()

# Let's summarize the paricipants by country
dat.per.country.2018 <- 
  dat.2018 %>% group_by(country) %>% 
  summarise(participants_2018 = n()) %>% arrange(desc(participants_2018)) %>% 
  # country names are in different formats;change them to a common one (iso3c format)
  mutate(country.iso3c = countrycode(country,origin='country.name', destination='iso3c'))

# Check for unknown country inputs
# dat.per.country.2018 %>% filter(is.na(country.iso3c)) # none found

# there is one participant from an unknown country; excluding them from the data
dat.per.country.2018 <- na.omit(dat.per.country.2018)

# Calculate the perecentages for country-wise participation
dat.per.country <- dat.per.country.2014 %>%
  left_join(dat.per.country.2018, by="country.iso3c") %>% 
  select(country.iso3c, participants_2014, participants_2018, everything()) %>% 
  mutate(Aus_2014 = round(participants_2014/sum(participants_2014)*100,1)) %>% 
  mutate(Bra_2018 = ifelse(is.na(participants_2018),
                               NA,
                               round(participants_2018/sum(na.omit(participants_2018))*100,1)))

# convert counrty iso codes to names for easy labeling in plots
dat.country.year <- 
  dat.per.country %>% 
  select(country.iso3c, Aus_2014, Bra_2018) %>% 
  pivot_longer(-country.iso3c, names_to = "year", values_to = "percentage") %>% 
  mutate(country.name = countrycode(country.iso3c, origin = 'iso3c', destination = 'country.name'))

# mean percent of participants from a country
mean.percent <- dat.country.year %>% pull(percentage) %>% na.omit() %>% mean()

```

___

### General Observations (data summary)

1. In 2014, there were `r sum(dat.per.country.2014$participants_2014)` participants from `r length(na.omit(dat.per.country.2014$country.iso3c))` countries, with a median of `r median(dat.per.country.2014$participants_2014)` participants/country.
2. In 2018, there were `r sum(dat.per.country.2018$participants_2018)` participants from `r length(na.omit(dat.per.country.2018$country.iso3c))` countries, with a median of `r median(dat.per.country.2018$participants_2018)` participants/country.

> Participants from 5 countries (USA, Australia, Japan, Germany and UK) account for `r round(sum(dat.per.country.2014$participants_2014[1:5])/sum(dat.per.country.2014$participants_2014)*100, 2)`% of all participants in IUSSI 2014 and `r round(sum(dat.per.country.2018$participants_2018[c(1,3:6)])/sum(dat.per.country.2018$participants_2018)*100, 2)`% in IUSSI 2018. 

___

### Participation by country

The following plot presents the percentage of IUSSI participants from different countries from the previous two IUSSIs (Aus_2014 = IUSSI 2014 in Australia, Bra_2018 = IUSSI 2018 in Brazil)

```{r, echo=F, fig.width=5,fig.height=10}
source(file = "./functions/theme_publication.R")
ggplot(dat.country.year,
       aes(x=reorder(country.name, percentage), 
           y=percentage, color=year)) +
  geom_hline(yintercept = mean.percent, color="grey60", alpha=0.6, lty=2) +
  geom_point(size=1.5, alpha=0.8) +
  
  labs(x = "",
       y = "Participants at IUSSI (%)") +
  coord_flip() +
  theme_Publication() +
  annotate("text", x = 4, y = 12, 
           label = paste0("mean participation \n per country (",round(mean.percent,1),"%)"))

```

___


In case you like tables or absolute counts over proportions, here is a searchable table that provides the number of participants, per country, that attended the previous two IUSSIs. 

```{r, echo=F}
library(DT)
participants.country.year <- 
  dat.per.country %>% select(country = country.x, 
                           IUSSI_2014 = participants_2014,
                           IUSSI_2018 = participants_2018)
DT::datatable(participants.country.year)
```

***

## Participation by Gender

Method: 

* I used genderizer.io for predicting gender from first names of the participants. The same tool is recommended by BiasWatchNeuro to estimate gender ratios in the absence of demographic data. The website for performing gender predictions can be found [here](https://genderize.io/check-gender-csv).

```{r, echo=F}
# Load the data
gen.dat.2014 <- read.csv("./data/genderized_data/genderized_2014_IUSSI.csv", header = T, stringsAsFactors = F)
gen.dat.2018 <- read.csv("./data/genderized_data/genderized_2018_IUSSI.csv", header = T, stringsAsFactors = F)

# define the cutoff for count per name in databse and probability of gender given name
count_cutoff <- 10
prob_cutoff <- 0.9

f.gen.dat.2014 <- 
  gen.dat.2014 %>% 
  select(country,
         gender, 
         count=genderCount,
         prob=genderProbability) %>% 
  filter(count > count_cutoff) %>% # 512/576 participants left at count_cutoff = 10
  arrange(count) %>% 
  filter(prob>prob_cutoff) %>%  # 461/576 participants left at count_cutoff = 10 and prob_cutoff = 0.9
  group_by(country, gender) %>% 
  summarize(participants_2014 = n()) %>% 
  mutate(country.iso3c = countrycode(country, origin='country.name', destination='iso3c'))
  
f.gen.dat.2018 <- 
  gen.dat.2018 %>% 
  select(country, gender,
         count=genderCount,
         prob=genderProbability) %>% 
  filter(count > count_cutoff & prob > prob_cutoff) %>% # 417/521 participants left after filtering
  arrange(count) %>% 
  group_by(country, gender) %>% 
  summarize(participants_2018 = n()) %>% 
  mutate(country.iso3c = countrycode(country, origin='country.name', destination='iso3c'))

```

* The results from genderizer.io were filtered to keep only participant names with a count greater than `r count_cutoff` in the database, and with probability of gender assignment higher than `r prob_cutoff`

* Post-filtering, a total of 461 (out of 576) participants in 2014 and 417 (out of 521) participants in 2018 remained. 
   

#### Estimates  

```{r, echo=F}
gender.ratio <-
  f.gen.dat.2014 %>%
  group_by(gender) %>%
  summarize(IUSSI_2014=sum(participants_2014)) %>%
  left_join((f.gen.dat.2018 %>%
              group_by(gender) %>%
              summarize(IUSSI_2018=sum(participants_2018))), by="gender")

library(kableExtra)
gender.ratio %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

> The, **estimated *female:male ratio* **among selected participants at **IUSSI 2014 was `r round(gender.ratio[1,2]/gender.ratio[2,2],2)`** and at **IUSSI 2018 was `r round(gender.ratio[1,3]/gender.ratio[2,3],2)`.**

***

## Participation by Career-stage

```{r, include=F}

## Exploring career stage of participants

titles.2014 <- dat.2014 %>% pull(title) %>% as.character()
titles.2018 <- dat.2018 %>% pull(degree) %>% as.character()

titles.2014 %>% unique()
titles.2018 %>% unique()

## Let's summarize participants by title (career-stage) 
  dat.2014 %>% 
  group_by(title) %>% 
  summarize(freq=n())
  
## Looks like it needs some clean-up
  # PIs - c("A/Prof","A/Professor","Ass/Prof","Associate Profes","Prof","Professor")
  # Graduate/Undergrads - c("Miss","Mr","Mrs","Ms")
  # Dr - most likely postdocs

career.2014 <- 
  dat.2014 %>% 
  mutate(degree = ifelse(title %in% c("A/Prof","A/Professor","Ass/Prof","Associate Profes","Prof","Professor"),
                         "Prof.",title)) %>% 
  mutate(degree = ifelse(title %in% c("Miss","Mr","Mrs","Ms"),
                         "Mr/Ms.",degree)) %>% 
  mutate(degree = ifelse(title %in% c("Dr"),
                         "Dr.",degree)) %>% 
  select(-title) %>% 
  select(degree, everything())

career.2014 <-
  career.2014 %>% 
  group_by(degree) %>% 
  summarise(freq = n()) %>% 
  filter(!degree %in% c("")) %>% # removed two with no assignment
  mutate(percent = round(freq/sum(freq)*100,1)) %>% 
  arrange(match(degree, c("Prof.","Dr.","Mr/Ms.")))


career.2018 <- 
  dat.2018 %>% 
  group_by(degree) %>% 
  summarize(freq=n()) %>% 
  filter(!degree %in% c("")) %>% # removed one with no assignment
  mutate(percent = round(freq/sum(freq)*100,1)) %>% 
  arrange(match(degree, c("Prof.","Dr.","Post-Doc","Graduate","Undergraduate","Other")))

```
   
* The data doesn't permit a comprehensive count of participants by their career stages since we have terms such as "Dr" that could indicate either a post-doc or a PI. 

* Additionally, the data for IUSSI 2014 has titles ("A/Prof","A/Professor","Ass/Prof","Associate Profes","Prof","Professor") that have been clubbed together as "Prof." and (Miss","Mr","Mrs","Ms") that are represented as "Mr/Ms."


```{r, echo=F}
career.2014 %>% 
  kable(caption = "IUSSI 2014") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

```{r, echo=F}
career.2018 %>% 
  kable(caption = "IUSSI 2018") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


> The data does show a **fairly good representation of graduate students** at both IUSSIs. 

However, in future, including the upcoming IUSSI 2022, it would be useful to come up with a standardized way to collect demography data during the registration process.

___

All code used for this document is available on my [Github](https://github.com/biplabendu/homepage/blob/master/IUSSI.Rmd).

___
