
---
title: "Demography data for IUSSI"
author: Biplabendu Das
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: yeti
    keep_md: no
    code_folding: hide
    includes:
      in_header: analytics.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

```

***

> What is this document?   

The goal:

- Perform a rapid exploratory analyses of the demography data obtained from the survey performed on the audience interested in or/and attending the upcoming IUSSI international conference in 2022, San Diego.

- sub-goals:
  - SUMMARIZE: who are these people?
  - ANALYZE 01: who are the folks that are underrepresented in attendence/presenting/speaking at IUSSI 2022?
  

```{r, include=F, echo=F}
# Let's load the data and get a glimpse.

rm(list = ls())
library(tidyverse)
library(kableExtra)

# Load custom functions
source("./functions/theme_publication.R")

raw.dat <- read.csv("./data/2022_demography_data/iussi_2022_data_28Jun22.csv", 
                header = T, stringsAsFactors = F,
                na.strings = c("NA",""," "))

# Get a glimpse
raw.dat %>% as.tibble() %>% head()

# Let's rename the columns to simple, easy-to-read/access data names
colnames(raw.dat)

new.colnames <- c("time_of_entry",
                  "age",
                  "gender",
                  "add_gender",
                  "sex_orientation",
                  "add_sex_orientation",
                  "residence_location",
                  "research_location",
                  "ethnic_identity",
                  "add_ethnic_identity",
                  "religious_identity",
                  "add_religious_identity",
                  "disability_identity",
                  "add_disability_identity",
                  "written_english",
                  "spoken_english",
                  "children_responsibility",
                  "adult_responsibility",
                  "funds_for_care_inhibit",
                  "firstgen",
                  "current_position",
                  "which_attended_before",
                  "role_attended_before",
                  "current_institute_desc",
                  "barriers_what_degree",
                  "barriers_which",
                  "add_barriers_which")

## Rename column names
colnames(raw.dat) <- new.colnames

```

***

## 01 Raw data

Let's take a look at few rows and columns of the raw data.

```{r raw_data, echo=F}
# get a glimpse of the column format features of the data
raw.dat[10:13,1:7] %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")

raw.dat[10:13,8:13] %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")

```

### What does each column contain?

**Column name and Description**

- time_of_entry: 
- age: 

***

## 02 Format data

> Get the data into a tidy format

- remove columns that are not useful for a first assessment
    - time_of_entry
    - add_xxx (all columns that contain additional response not listed)
- format each column by
    - separating multiple choices into separate rows
    - factorizing the variables
    - checking for erronneous choices/inputs


```{r format_data}

## Tidy and save data
dat <- 
  
  raw.dat %>% as_tibble() %>% 
  
  
  ## remove columns that you do not want
  select(-time_of_entry) %>% 
  select(-starts_with("add")) %>% 
  
  ## add a unique ID for each person
  rownames_to_column("ID") %>% 
  
  ## AGE
  mutate(age = str_split_fixed(age, " ", n=2)[,1]) %>%
  mutate(age = ifelse(age %in% c("","Prefer"), "Other", age)) %>%
  mutate(age = as.factor(age)) %>%
  # ## Plot: AGE
  # group_by(age) %>%
  # summarise(num = n()) %>% 
  # ##
  # ggplot(aes(x=age, y=num)) +
  #   geom_point() +
  #   coord_flip() +
  #   theme_Publication()

  
  ## GENDER
    # there are two entries that contains the following two responses for gender identity
    # c("Cisgender woman;Prefer not to respond", "Transgender man;Cisgender woman")),
    # five entries that stated "Prefer not to respond", and one which is NA.
    # Rename all to "Other"
  mutate(gender = ifelse(!gender %in% c("Cisgender man", "Cisgender woman", "Non-binary",
                       "Transgender man", "Transgender woman"), "Other", gender)) %>%
  mutate(gender = as.factor(gender)) %>%
      # ## Plot: GENDER
      # group_by(gender) %>%
      # summarise(num = n()) %>%
      # na.omit() %>%
      # ##
      # ggplot(aes(x=gender, y=num/sum(num)*100)) +
      #   geom_point() +
      #   coord_flip() +
      #   theme_Publication()

  
  ## sex_orientation
  mutate(sex_orientation = ifelse(sex_orientation %in% c("Additional identity not listed? (Please click next and share)","Prefer not to respond"), "Other", sex_orientation)) %>% 
  mutate(sex_orientation = as.factor(sex_orientation)) %>% 
  
  
  ## residence_location
  mutate(residence_location = ifelse(residence_location %in% c("Central America;Other",
                                     "Continental US;South Asia",
                                     "Europe;Middle East",
                                     "Europe;Southeast Asia",
                                     "Other"), "Other", residence_location)) %>%
  mutate(residence_location = as.factor(residence_location)) %>% 
      # ## Plot: residence_location + gender
      # group_by(residence_location, gender) %>% 
      # summarize(num = n()) %>% 
      # ggplot(aes(x=residence_location, y=log2(num+1), fill=gender, group=gender)) +
      # # geom_point(size=4, alpha=0.85) +
      # geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
      # geom_bar(position="stack", stat="identity") +
      # coord_flip() +
      # theme_Publication() +
      # scale_colour_Publication() +
      # scale_fill_Publication() +
      # guides(fill = guide_legend(nrow = 3))


  ## research location
  # filter(research_location %in% c("Central America;Continental US", "Europe;South America")) %>%
  separate_rows(research_location, sep = ";") %>% 
  mutate(research_location = as.factor(research_location)) %>% 
      # ## Plot: research_location + gender
      # group_by(research_location, gender) %>% 
      # summarize(num = n()) %>% 
      # ggplot(aes(x=research_location, y=log2(num+1), fill=gender, group=gender)) +
      # geom_bar(position="stack", stat="identity", size=2) +
      # # geom_point(size=4, alpha=0.85) +
      # geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
      # coord_flip() +
      # theme_Publication() +
      # scale_colour_Publication() +
      # scale_fill_Publication() +
      # guides(fill = guide_legend(nrow = 3))
  
  
  ## ethnic_identity
  separate_rows(ethnic_identity, sep = ";") %>% 
  mutate(ethnic_identity = ifelse(ethnic_identity %in% c("A race or ethnicity not listed",
                                                         "Prefer not to respond"), 
                                  "Other", ethnic_identity)) %>% 
  mutate(ethnic_identity = as.factor(ethnic_identity)) %>% 
  
  
  ## religious_identity
  mutate(religious_identity = ifelse(religious_identity %in% c("An additional affiliation not listed",
                                                               "Prefer not to respond"), 
                                     "Other", religious_identity)) %>% 
  mutate(religious_identity = as.factor(religious_identity)) %>% 
  
  
  ## disability_identity
  separate_rows(disability_identity, sep = ";") %>% 
  mutate(disability_identity = as.factor(disability_identity)) %>% 
  

  ## written_english
  mutate(written_english = as.factor(written_english)) %>% 
  
  
  ## spoken_english
  mutate(spoken_english = as.factor(spoken_english)) %>% 
  
  
  ## children_responsibility
  mutate(children_responsibility = as.factor(children_responsibility)) %>% 
  
  
  ## adult_responsibility
  mutate(adult_responsibility = ifelse(adult_responsibility %in% c("Prefer not to respond", "Other"),
                                       "Other", adult_responsibility)) %>% 
  mutate(adult_responsibility = as.factor(adult_responsibility)) %>% 
  
  
  ## funds_for_care_inhibit
  mutate(funds_for_care_inhibit = as.factor(funds_for_care_inhibit)) %>% 
  
  
  ## firstgen
  mutate(firstgen = as.factor(firstgen)) %>%  
  
  
  ## current position
  separate_rows(current_position, sep = ";") %>% 
  mutate(current_position = as.factor(current_position)) %>% 
    # ## Plot: research_location + gender
    # group_by(current_position, gender) %>%
    # summarize(num = n()) %>%
    #   ggplot(aes(x=current_position, y=log2(num+1), fill=gender, group=gender)) +
    #   geom_bar(position="stack", stat="identity", size=2) +
    #   # geom_point(size=4, alpha=0.85) +
    #   geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
    #   coord_flip() +
    #   theme_Publication() +
    #   scale_colour_Publication() +
    #   scale_fill_Publication() +
    #   guides(fill = guide_legend(nrow = 3))
    
  
  ## which_attended_before
  separate_rows(which_attended_before, sep = ";") %>% 
  mutate(which_attended_before = as.factor(which_attended_before)) %>% 

  
  ## role_attended_before
  separate_rows(role_attended_before, sep = ";") %>% 
  mutate(role_attended_before = as.factor(role_attended_before)) %>% 
    # ## Plot: research_location + gender
    # group_by(role_attended_before, gender) %>%
    # summarize(num = n()) %>%
    #   ggplot(aes(x=role_attended_before, y=log2(num+1), fill=gender, group=gender)) +
    #   geom_bar(position="stack", stat="identity", size=2) +
    #   # geom_point(size=4, alpha=0.85) +
    #   geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
    #   coord_flip() +
    #   theme_Publication() +
    #   scale_colour_Publication() +
    #   scale_fill_Publication() +
    #   guides(fill = guide_legend(nrow = 3))

  
  ## current_institute_desc
  separate_rows(current_institute_desc, sep = ";") %>% 
  mutate(current_institute_desc = as.factor(current_institute_desc)) %>% 
  
  
  ## barriers_what_degree
  mutate(barriers_what_degree = as.factor(barriers_what_degree)) %>% 
  
  
  ## barriers_which
  separate_rows(barriers_which, sep = ";") %>% 
  mutate(barriers_which = as.factor(barriers_which))


# ## get a glimpse of the tidy data
# dat[10:13,1:7] %>% 
#   kable() %>% 
#   kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")
# 
# dat[10:13,8:13] %>% 
#   kable() %>% 
#   kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")

```


***

## 03 Visualize

### AGE & GENDER

```{r age_gender}

## AGE + GENDER
dat %>% 
  
  # remove the rows with no age/gender info
  filter(age != "Other") %>% 
  filter(gender != "Other") %>% 
  
  # tally the number of people in a given age group for a given gender
  distinct(ID, age, gender) %>%
  group_by(age, gender) %>% 
  summarize(num=n()) %>% 
  na.omit() %>%
    
  ## Plot: AGE + GENDER
  ggplot(aes(x=age, y=num, col=gender, group=gender)) +
    labs(y="number of people") +
    geom_point() +
    geom_line() +
    coord_flip() +
    theme_Publication(base_size = 16) +
    scale_colour_Publication() +
    guides(color = guide_legend(nrow = 3))

```



___

All code used for this document is available on my [Github](https://github.com/biplabendu/homepage/blob/master/IUSSI.Rmd).

___
