
---
title: "Demography data for IUSSI"
author:
- name: Biplabendu Das
  affiliation: DEIJ Committee, IUSSI
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: yeti
    keep_md: no
    code_folding: hide
    includes:
      in_header: analytics.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

```

***

> What is this document?   

The goal:

Perform a rapid exploratory analyses of the demography data obtained from the survey performed on the audience interested in or/and attending the upcoming IUSSI international conference in 2022, San Diego.



```{r, include=F, echo=F}
# Let's load the data and get a glimpse.

rm(list = ls())
library(tidyverse)
library(kableExtra)

# Load custom functions
source("./functions/theme_publication.R")

raw.dat <- read.csv("./data/2022_demography_data/iussi_2022_data_28Jun22.csv", 
                header = T, stringsAsFactors = F,
                na.strings = c("NA",""," "))

# Get a glimpse
raw.dat %>% as.tibble() %>% head()

# Let's rename the columns to simple, easy-to-read/access data names
colnames(raw.dat)

new.colnames <- c("time_of_entry",
                  "age",
                  "gender",
                  "add_gender",
                  "sex_orientation",
                  "add_sex_orientation",
                  "residence_location",
                  "research_location",
                  "ethnic_identity",
                  "add_ethnic_identity",
                  "religious_identity",
                  "add_religious_identity",
                  "disability_identity",
                  "add_disability_identity",
                  "written_english",
                  "spoken_english",
                  "children_responsibility",
                  "adult_responsibility",
                  "funds_for_care_inhibit",
                  "firstgen",
                  "current_position",
                  "which_attended_before",
                  "role_attended_before",
                  "current_institute_desc",
                  "barriers_what_degree",
                  "barriers_which",
                  "add_barriers_which")

## Rename column names
colnames(raw.dat) <- new.colnames

```

***

## 01 Raw data

Let's take a look at few rows and columns of the raw data.

```{r raw_data, echo=F}
# get a glimpse of the column format features of the data
raw.dat[10:13,1:7] %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")

raw.dat[10:13,8:13] %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")

```

### What does each column contain?

**Column name: Description**

- time_of_entry: 
- age: 
- and so on..

***

## 02 Format data

> Get the data into a tidy format

- remove columns that are not useful for a first assessment
    - time_of_entry
    - add_xxx (all columns that contain additional response not listed)
- format each column by
    - separating multiple choices into separate rows
    - factorizing the variables
    - checking for erronneous choices/inputs


```{r format_data}

## Tidy and save data
dat <- 
  
  raw.dat %>% as_tibble() %>% 
  
  
  ## remove columns that you do not want
  select(-time_of_entry) %>% 
  select(-starts_with("add")) %>% 
  
  ## add a unique ID for each person
  rownames_to_column("ID") %>% 
  
  ## AGE
  mutate(age = str_split_fixed(age, " ", n=2)[,1]) %>%
  mutate(age = ifelse(age %in% c("","Prefer"), "Other", age)) %>%
  mutate(age = as.factor(age)) %>%
      # ## Plot: AGE
      # group_by(age) %>%
      # summarise(num = n()) %>% 
      # ##
      # ggplot(aes(x=age, y=num)) +
      #   geom_point() +
      #   coord_flip() +
      #   theme_Publication()

  
  ## GENDER
    # there are two entries that contains the following two responses for gender identity
    # c("Cisgender woman;Prefer not to respond", "Transgender man;Cisgender woman")),
    # five entries that stated "Prefer not to respond", and one which is NA.
    # Rename all to "Other"
  mutate(gender = ifelse(!gender %in% c("Cisgender man", "Cisgender woman", "Non-binary",
                       "Transgender man", "Transgender woman"), "Other", gender)) %>%
  mutate(gender = as.factor(gender)) %>%
      # ## Plot: GENDER
      # group_by(gender) %>%
      # summarise(num = n()) %>%
      # na.omit() %>%
      # ##
      # ggplot(aes(x=gender, y=num/sum(num)*100)) +
      #   geom_point() +
      #   coord_flip() +
      #   theme_Publication()

  
  ## sex_orientation
  mutate(sex_orientation = ifelse(sex_orientation %in% c("Additional identity not listed? (Please click next and share)","Prefer not to respond"), "Other", sex_orientation)) %>% 
  mutate(sex_orientation = as.factor(sex_orientation)) %>% 
  
  
  ## residence_location
  mutate(residence_location = ifelse(residence_location %in% c("Central America;Other",
                                     "Continental US;South Asia",
                                     "Europe;Middle East",
                                     "Europe;Southeast Asia",
                                     "Other"), "Other", residence_location)) %>%
  mutate(residence_location = as.factor(residence_location)) %>% 


  ## research location
  # filter(research_location %in% c("Central America;Continental US", "Europe;South America")) %>%
  separate_rows(research_location, sep = ";") %>% 
  mutate(research_location = as.factor(research_location)) %>% 
  
  
  ## ethnic_identity
  separate_rows(ethnic_identity, sep = ";") %>% 
  mutate(ethnic_identity = ifelse(ethnic_identity %in% c("A race or ethnicity not listed",
                                                         "Prefer not to respond"), 
                                  "Other", ethnic_identity)) %>% 
  mutate(ethnic_identity = as.factor(ethnic_identity)) %>% 
  
  
  ## religious_identity
  mutate(religious_identity = ifelse(religious_identity %in% c("An additional affiliation not listed",
                                                               "Prefer not to respond"), 
                                     "Other", religious_identity)) %>% 
  mutate(religious_identity = as.factor(religious_identity)) %>% 
  
  
  ## disability_identity
  separate_rows(disability_identity, sep = ";") %>% 
  mutate(disability_identity = as.factor(disability_identity)) %>% 
  

  ## written_english
  mutate(written_english = as.factor(written_english)) %>% 
  
  
  ## spoken_english
  mutate(spoken_english = as.factor(spoken_english)) %>% 
  
  
  ## children_responsibility
  mutate(children_responsibility = as.factor(children_responsibility)) %>% 
  
  
  ## adult_responsibility
  mutate(adult_responsibility = ifelse(adult_responsibility %in% c("Prefer not to respond", "Other"),
                                       "Other", adult_responsibility)) %>% 
  mutate(adult_responsibility = as.factor(adult_responsibility)) %>% 
  
  
  ## funds_for_care_inhibit
  mutate(funds_for_care_inhibit = as.factor(funds_for_care_inhibit)) %>% 
  
  
  ## firstgen
  mutate(firstgen = as.factor(firstgen)) %>%  
  
  
  ## current position
  separate_rows(current_position, sep = ";") %>% 
  mutate(current_position = as.factor(current_position)) %>% 
  
  
  ## which_attended_before
  separate_rows(which_attended_before, sep = ";") %>% 
  mutate(which_attended_before = as.factor(which_attended_before)) %>% 

  
  ## role_attended_before
  separate_rows(role_attended_before, sep = ";") %>% 
  mutate(role_attended_before = as.factor(role_attended_before)) %>% 

  
  ## current_institute_desc
  separate_rows(current_institute_desc, sep = ";") %>% 
  mutate(current_institute_desc = as.factor(current_institute_desc)) %>% 
  
  
  ## barriers_what_degree
  mutate(barriers_what_degree = as.factor(barriers_what_degree)) %>% 
  
  
  ## barriers_which
  separate_rows(barriers_which, sep = ";") %>% 
  mutate(barriers_which = as.factor(barriers_which))


# ## get a glimpse of the tidy data
# dat[10:13,1:7] %>% 
#   kable() %>% 
#   kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")
# 
# dat[10:13,8:13] %>% 
#   kable() %>% 
#   kable_styling(bootstrap_options = c("striped"), full_width = F, position = "center", latex_options = "scaled_down")

```


***

## 03 Visualize

```{r define_params}

## Use this code chunk to define paramaters, such as color palettes and such

## COLOR
# role attended before
roles.color <- c("#005C53", "#F2E744", "#D9814E", "#BF6854", "#261B1A", "grey60")

```



### Age & gender

```{r age_gender}

## AGE + GENDER
dat %>% 
  
  # remove the rows with no age/gender info
  filter(age != "Other") %>% 
  filter(!is.na(gender)) %>%
  # filter(gender != "Other") %>% 
  
  # tally the number of people in a given age group for a given gender
  distinct(ID, age, gender) %>%
  group_by(age, gender) %>% 
  summarize(num=n()) %>% 
    
  ## Plot: AGE + GENDER
  ggplot(aes(x=age, y=num, fill=gender, group=gender)) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    geom_bar(position="stack", stat="identity") +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_Publication() +
    guides(fill = guide_legend(nrow = 3))

```

***

### Residence location & gender

```{r residence_location_gender}
# Residence location & gender
dat %>%

  # filter(!residence_location %in% c("Other")) %>% 
  filter(!is.na(residence_location)) %>%
  filter(!is.na(gender)) %>%
  # filter(gender != "Other") %>% 
  
  distinct(ID, residence_location, gender) %>% 
  group_by(residence_location, gender) %>%
  summarize(num = n()) %>% 
  
  ggplot(aes(x=residence_location, y=num, fill=gender, group=gender)) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    geom_bar(position="stack", stat="identity") +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_Publication() +
    guides(fill = guide_legend(nrow = 3))

```

***

### Research location & gender

```{r research_location_gender}

## Plot: research_location + gender
dat %>% 
  
  filter(!is.na(research_location)) %>%
  filter(!is.na(gender)) %>%
  # filter(gender != "Other") %>%
  
  distinct(ID, research_location, gender) %>% 
  group_by(research_location, gender) %>%
  summarize(num = n()) %>% 
  
  ggplot(aes(x=research_location, y=num, fill=gender, group=gender)) +
    geom_bar(position="stack", stat="identity", size=2) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_Publication() +
    guides(fill = guide_legend(nrow = 3))
```

> A bunch of folks perform research in Central America, but there seems to be no representation from folks living in Central America. There is, of course, the possibility that folks residing in Central America did not take the survey.

***

### Role of attendee & gender

```{r role_attended_before_gender}

dat %>% 
  
  # remove the rows with no info
  filter(!is.na(role_attended_before)) %>%
  filter(!is.na(gender)) %>%
  # filter(gender != "Other") %>% 
  
  distinct(ID, role_attended_before, gender) %>% 
  group_by(role_attended_before, gender) %>%
  summarize(num = n()) %>% 
    
  ## Plot: research_location + gender
  ggplot(aes(x=role_attended_before, y=num, fill=gender, group=gender)) +
    geom_bar(position="stack", stat="identity", size=2) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_Publication() +
    guides(fill = guide_legend(nrow = 3))

```

***

### Current position & gender

```{r current_position_gender}

dat %>% 
  
  filter(!is.na(current_position)) %>%
  filter(!is.na(gender)) %>%
  # filter(gender != "Other") %>% 

  distinct(ID, current_position, gender) %>% 
  group_by(current_position, gender) %>%
  summarize(num = n()) %>% 
  
  ggplot(aes(x=current_position, y=log2(num+1), fill=gender, group=gender)) +
    geom_bar(position="stack", stat="identity", size=2) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_Publication() +
    guides(fill = guide_legend(nrow = 3))

```

***
***

### Role of attendee & residence location

```{r role_attended_before_residence_location}

dat %>% 
  
  # remove the rows with no age/gender info
  filter(!is.na(role_attended_before)) %>%
  filter(!is.na(residence_location)) %>%
  # filter(gender != "Other") %>% 
  
  distinct(ID, role_attended_before, residence_location) %>% 
  group_by(role_attended_before, residence_location) %>%
  summarize(num = n()) %>%
    
  ggplot(aes(x=residence_location, y=num, fill=role_attended_before, group=role_attended_before)) +
    geom_bar(position="stack", stat="identity", size=2) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    # scale_fill_Publication() +
    scale_fill_manual(values = roles.color) +
    guides(fill = guide_legend(nrow = 3))

```

***

### Role of attendee & current position

```{r role_attended_before_current_position}

dat %>% 
  
  # remove the rows with no age/gender info
  filter(!is.na(role_attended_before)) %>%
  filter(!is.na(current_position)) %>%
  # filter(gender != "Other") %>% 
  
  distinct(ID, role_attended_before, current_position) %>% 
  group_by(role_attended_before, current_position) %>%
  summarize(num = n()) %>% 
    
  ## Plot: research_location + gender
  ggplot(aes(x=current_position, y=num, fill=role_attended_before, group=role_attended_before)) +
    geom_bar(position="stack", stat="identity", size=2) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_manual(values = roles.color) +
    guides(fill = guide_legend(nrow = 3))
# to do:
# normalize the number of roles undertaken by a person in past IUSSI
# number of past roles / number of past IUSSI attended

```

The above plot might be misleading since the higher number of roles assumed by (activity of) late, tenured Profs might be due to the higher number of previous conferences attened by them. So, it might make sense to have a standardized metric that gives us a better understanding of the activity of each group in IUSSI.

> Activity = number of roles assumed (other than attendee) / number of IUSSIs attended


```{r role_attended_before_current_position_v2}

id <- character()
num_attended <- numeric()  
for (i in 1:length(levels(as.factor(dat$ID)))) {
    
    id[i] <- levels(as.factor(dat$ID))[i]
    
    num_attended[i] <-
      dat %>% 
      filter(ID == id[i]) %>%
      # select(which_attended_before)
      mutate(which_attended_before=as.character(which_attended_before)) %>% 
      pull(which_attended_before) %>% 
      unique() %>% 
      as.factor() %>% 
      levels() %>% length() %>% 
      as.numeric()
}

dummy.dat <- data.frame(ID=id, num_attended=num_attended)

## Add this information to the primary dataset

dat %>% 
  
  # add the num_attended column to dat
  left_join(dummy.dat, by="ID") %>% 
    # check to see if it worked
    # select(ID, which_attended_before, num_attended) %>% head(20)
  
  # remove the rows with no info
  filter(!is.na(role_attended_before)) %>%
  filter(!is.na(current_position)) %>%
  # filter(gender != "Other") %>% 
  
  select(ID, current_position, role_attended_before, num_attended) %>% 
    # head(20)
  
  distinct(ID, role_attended_before, current_position, num_attended) %>% 
  
  group_by(ID, current_position, num_attended) %>%
  summarize(num_roles = n()) %>% 
  arrange(as.numeric(ID)) %>% 
  
  # calculate activity
  mutate(activity = num_roles/num_attended) %>% 
    
  ## Plot: research_location + gender
  ggplot(aes(x=current_position, y=activity,fill=current_position)) +
    # geom_bar(position="stack", stat="identity", size=2) +
    geom_boxplot(alpha=0.7, outlier.colour = "darkred") +
    labs(y="roles/attendance") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_Publication() +
    # scale_fill_manual(values = roles.color) +
    guides(fill = guide_legend(nrow = 8))
```

***

### Role of attendee & ethnic identity

```{r role_attended_before_ethnic_identity}

dat %>% 
  
  # remove the rows with no age/gender info
  filter(!is.na(role_attended_before)) %>%
  filter(!is.na(ethnic_identity)) %>%
  # filter(gender != "Other") %>% 
  
  distinct(ID, role_attended_before, ethnic_identity) %>% 
  group_by(role_attended_before, ethnic_identity) %>%
  summarize(num = n()) %>% 
    
  ## Plot: research_location + gender
  ggplot(aes(x=ethnic_identity, y=num, fill=role_attended_before, group=role_attended_before)) +
    geom_bar(position="stack", stat="identity", size=2) +
    labs(y="number of people") +
    # geom_hline(yintercept = 5, col="grey60", alpha=0.7, size=1) +
    coord_flip() +
    theme_Publication() +
    scale_colour_Publication() +
    scale_fill_manual(values = roles.color) +
    guides(fill = guide_legend(nrow = 3))

```

___

All code used for this document is available on my [Github](https://github.com/biplabendu/homepage/blob/master/IUSSI2022.Rmd).

___

## 99. IUSSI 2014 & 2018

The demography data for the participants that attended IUSSI 2014 and 2018 can be found [here](https://biplabendu.github.io/homepage/IUSSI.html).