
---
title: "Who is attending IUSSI 2022?"
author: Biplabendu Das
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    theme: yeti
    includes:
      in_header: analytics.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

```

***

> What is this document?   

The goal:

- Perform a rapid exploratory analyses of the demography data obtained from the survey performed on the audience interested in or/and attending the upcoming IUSSI international conference in 2022, San Diego.

- sub-goals:
  - SUMMARIZE: who are these people?
  - ANALYZE 01: who are the folks that are underrepresented in attendence/presenting/speaking at IUSSI 2022?
  

```{r, include=F, echo=F}
# Let's load the data and get a glimpse.

rm(list = ls())
library(tidyverse)
library(countrycode)

# Load custom functions
source("./functions/theme_publication.R")

dat <- read.csv("./data/2022_demography_data/iussi_2022_data_28Jun22.csv", 
                header = T, stringsAsFactors = F,
                na.strings = c("NA",""," "))

# Get a glimpse
dat %>% as.tibble() %>% head()

# Let's rename the columns to simple, easy-to-read/access data names
colnames(dat)

new.colnames <- c("time_of_entry",
                  "age",
                  "gender",
                  "add_gender",
                  "sex_orientation",
                  "add_sex_orientation",
                  "residence_location",
                  "research_location",
                  "ethnic_identity",
                  "add_ethnic_identity",
                  "religious_identity",
                  "add_religious_identity",
                  "disability_identity",
                  "add_disability_identity",
                  "written_english",
                  "spoken_english",
                  "children_responsibility",
                  "adult_responsibility",
                  "funds_for_care_inhibit",
                  "firstgen",
                  "current_position",
                  "which_attended_before",
                  "role_attended_before",
                  "current_institute_desc",
                  "barriers_what_degree",
                  "barriers_which",
                  "add_barriers_which")

## Rename column names
colnames(dat) <- new.colnames

# get a glimpse of the column format features of the data
dat %>% glimpse()

## Assign proper data formats
dat %>% 
  
  # remove columns that you do not want
  select(-time_of_entry) %>% 
  
  ## AGE
  mutate(age = str_split_fixed(age, " ", n=2)[,1]) %>%
  mutate(age = as.factor(age)) %>%
  filter(!age %in% c("","Prefer")) %>%
  # ## Plot: AGE
  # group_by(age) %>%
  # summarise(num = n()) %>% 
  # ##
  # ggplot(aes(x=age, y=num)) +
  #   geom_point() +
  #   coord_flip() +
  #   theme_Publication()

  ## GENDER
  mutate(gender = as.factor(gender)) %>% 
  filter(gender %in% c("Cisgender man", "Cisgender woman", "Non-binary",
                       "Transgender man", "Transgender woman")) %>%
      # ## Plot: GENDER
      # group_by(gender) %>%
      # summarise(num = n()) %>%
      # na.omit() %>%
      # ##
      # ggplot(aes(x=gender, y=num/sum(num)*100)) +
      #   geom_point() +
      #   coord_flip() +
      #   theme_Publication()
  
  # ## AGE + GENDER
  # group_by(age, gender) %>% 
  # summarize(num=n()) %>% 
  # na.omit() %>% 
      # ## Plot: AGE + GENDER
      # ggplot(aes(x=age, y=num, col=gender, group=gender)) +
      #   geom_point() +
      #   geom_line() +
      #   coord_flip() +
      #   theme_Publication() +
      #   scale_colour_Publication() +
      #   guides(color = guide_legend(nrow = 3))
  
  ## residence_location
  mutate(residence_location = as.factor(residence_location)) %>% 
      # ## Plot: residence_location + gender
      # group_by(residence_location, gender) %>% 
      # summarize(num = n()) %>% 
      # ggplot(aes(x=residence_location, y=log2(num+1), fill=gender, group=gender)) +
      # # geom_point(size=4, alpha=0.85) +
      # geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
      # geom_bar(position="stack", stat="identity") +
      # coord_flip() +
      # theme_Publication() +
      # scale_colour_Publication() +
      # scale_fill_Publication() +
      # guides(fill = guide_legend(nrow = 3))

  ## research location
  mutate(research_location = as.factor(research_location)) %>% 
      # ## Plot: research_location + gender
      # group_by(research_location, gender) %>% 
      # summarize(num = n()) %>% 
      # ggplot(aes(x=research_location, y=log2(num+1), fill=gender, group=gender)) +
      # geom_bar(position="stack", stat="identity", size=2) +
      # # geom_point(size=4, alpha=0.85) +
      # geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
      # coord_flip() +
      # theme_Publication() +
      # scale_colour_Publication() +
      # scale_fill_Publication() +
      # guides(fill = guide_legend(nrow = 3))

  ## current position
  mutate(current_position = as.factor(current_position)) %>% 
    # ## Plot: research_location + gender
    # group_by(current_position, gender) %>%
    # summarize(num = n()) %>%
    #   ggplot(aes(x=current_position, y=log2(num+1), fill=gender, group=gender)) +
    #   geom_bar(position="stack", stat="identity", size=2) +
    #   # geom_point(size=4, alpha=0.85) +
    #   geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
    #   coord_flip() +
    #   theme_Publication() +
    #   scale_colour_Publication() +
    #   scale_fill_Publication() +
    #   guides(fill = guide_legend(nrow = 3))

  ## role_attended_before
  mutate(role_attended_before = as.factor(role_attended_before)) %>% 
    ## Plot: research_location + gender
    group_by(role_attended_before, gender) %>%
    summarize(num = n()) %>%
      ggplot(aes(x=role_attended_before, y=log2(num+1), fill=gender, group=gender)) +
      geom_bar(position="stack", stat="identity", size=2) +
      # geom_point(size=4, alpha=0.85) +
      geom_hline(yintercept = log2(2), col="grey60", alpha=0.5, size=2) +
      coord_flip() +
      theme_Publication() +
      scale_colour_Publication() +
      scale_fill_Publication() +
      guides(fill = guide_legend(nrow = 3))

  as_tibble() %>% 
  # select(age, age_low, age_up) %>%
  select(age, gender, residence_location, current_position) %>%
  head()


# # Let's summarize the paricipants by country
# dat.per.country.2014 <- 
#   dat.2014 %>% group_by(country) %>% 
#   summarise(participants_2014 = n()) %>% arrange(desc(participants_2014)) %>% 
#   # country names are in different formats;change them to a common one (iso3c format)
#   mutate(country.iso3c = countrycode(country,origin='country.name', destination='iso3c'))
# 
# # How many countries were present?
# dat.per.country.2014 %>% nrow()
# 
# # Any unknown countries?
# # dat.per.country.2014 %>% filter(is.na(country.iso3c)) # nope
# 
# # What's the median number of participants per country?
# dat.per.country.2014 %>% pull(participants_2014) %>% median()
# 
# # Let's take a look at the top 15 and bottom 15 countries by participation
# dat.per.country.2014 %>% head(15)
# dat.per.country.2014 %>% tail(15)


```



```{r, inlcude=F, echo=F}
## Let's compare 2014 and 2018
dat.2018 <- read.csv("./data/2018_participants_list.csv", header = T, stringsAsFactors = F)

# # Get a glimpse
# dat.2018 %>% glimpse()

# Let's summarize the paricipants by country
dat.per.country.2018 <- 
  dat.2018 %>% group_by(country) %>% 
  summarise(participants_2018 = n()) %>% arrange(desc(participants_2018)) %>% 
  # country names are in different formats;change them to a common one (iso3c format)
  mutate(country.iso3c = countrycode(country,origin='country.name', destination='iso3c'))

# Check for unknown country inputs
# dat.per.country.2018 %>% filter(is.na(country.iso3c)) # none found

# there is one participant from an unknown country; excluding them from the data
dat.per.country.2018 <- na.omit(dat.per.country.2018)

# Calculate the perecentages for country-wise participation
dat.per.country <- dat.per.country.2014 %>%
  left_join(dat.per.country.2018, by="country.iso3c") %>% 
  select(country.iso3c, participants_2014, participants_2018, everything()) %>% 
  mutate(Aus_2014 = round(participants_2014/sum(participants_2014)*100,1)) %>% 
  mutate(Bra_2018 = ifelse(is.na(participants_2018),
                               NA,
                               round(participants_2018/sum(na.omit(participants_2018))*100,1)))

# convert counrty iso codes to names for easy labeling in plots
dat.country.year <- 
  dat.per.country %>% 
  select(country.iso3c, Aus_2014, Bra_2018) %>% 
  pivot_longer(-country.iso3c, names_to = "year", values_to = "percentage") %>% 
  mutate(country.name = countrycode(country.iso3c, origin = 'iso3c', destination = 'country.name'))

# mean percent of participants from a country
mean.percent <- dat.country.year %>% pull(percentage) %>% na.omit() %>% mean()

```

___

### General Observations (data summary)

1. In 2014, there were `r sum(dat.per.country.2014$participants_2014)` participants from `r length(na.omit(dat.per.country.2014$country.iso3c))` countries, with a median of `r median(dat.per.country.2014$participants_2014)` participants/country.
2. In 2018, there were `r sum(dat.per.country.2018$participants_2018)` participants from `r length(na.omit(dat.per.country.2018$country.iso3c))` countries, with a median of `r median(dat.per.country.2018$participants_2018)` participants/country.

> Participants from 5 countries (USA, Australia, Japan, Germany and UK) account for `r round(sum(dat.per.country.2014$participants_2014[1:5])/sum(dat.per.country.2014$participants_2014)*100, 2)`% of all participants in IUSSI 2014 and `r round(sum(dat.per.country.2018$participants_2018[c(1,3:6)])/sum(dat.per.country.2018$participants_2018)*100, 2)`% in IUSSI 2018. 

___

### Participation by country

The following plot presents the percentage of IUSSI participants from different countries from the previous two IUSSIs (Aus_2014 = IUSSI 2014 in Australia, Bra_2018 = IUSSI 2018 in Brazil)

```{r, echo=F, fig.width=5,fig.height=10}
source(file = "./functions/theme_publication.R")
ggplot(dat.country.year,
       aes(x=reorder(country.name, percentage), 
           y=percentage, color=year)) +
  geom_hline(yintercept = mean.percent, color="grey60", alpha=0.6, lty=2) +
  geom_point(size=1.5, alpha=0.8) +
  
  labs(x = "",
       y = "Participants at IUSSI (%)") +
  coord_flip() +
  theme_Publication() +
  annotate("text", x = 4, y = 12, 
           label = paste0("mean participation \n per country (",round(mean.percent,1),"%)"))

```

___


In case you like tables or absolute counts over proportions, here is a searchable table that provides the number of participants, per country, that attended the previous two IUSSIs. 

```{r, echo=F}
library(DT)
participants.country.year <- 
  dat.per.country %>% select(country = country.x, 
                           IUSSI_2014 = participants_2014,
                           IUSSI_2018 = participants_2018)
DT::datatable(participants.country.year)
```

***

## Participation by Gender

Method: 

* I used genderizer.io for predicting gender from first names of the participants. The same tool is recommended by BiasWatchNeuro to estimate gender ratios in the absence of demographic data. The website for performing gender predictions can be found [here](https://genderize.io/check-gender-csv).

```{r, echo=F}
# Load the data
gen.dat.2014 <- read.csv("./data/genderized_data/genderized_2014_IUSSI.csv", header = T, stringsAsFactors = F)
gen.dat.2018 <- read.csv("./data/genderized_data/genderized_2018_IUSSI.csv", header = T, stringsAsFactors = F)

# define the cutoff for count per name in databse and probability of gender given name
count_cutoff <- 10
prob_cutoff <- 0.9

f.gen.dat.2014 <- 
  gen.dat.2014 %>% 
  select(country,
         gender, 
         count=genderCount,
         prob=genderProbability) %>% 
  filter(count > count_cutoff) %>% # 512/576 participants left at count_cutoff = 10
  arrange(count) %>% 
  filter(prob>prob_cutoff) %>%  # 461/576 participants left at count_cutoff = 10 and prob_cutoff = 0.9
  group_by(country, gender) %>% 
  summarize(participants_2014 = n()) %>% 
  mutate(country.iso3c = countrycode(country, origin='country.name', destination='iso3c'))
  
f.gen.dat.2018 <- 
  gen.dat.2018 %>% 
  select(country, gender,
         count=genderCount,
         prob=genderProbability) %>% 
  filter(count > count_cutoff & prob > prob_cutoff) %>% # 417/521 participants left after filtering
  arrange(count) %>% 
  group_by(country, gender) %>% 
  summarize(participants_2018 = n()) %>% 
  mutate(country.iso3c = countrycode(country, origin='country.name', destination='iso3c'))

```

* The results from genderizer.io were filtered to keep only participant names with a count greater than `r count_cutoff` in the database, and with probability of gender assignment higher than `r prob_cutoff`

* Post-filtering, a total of 461 (out of 576) participants in 2014 and 417 (out of 521) participants in 2018 remained. 
   

#### Estimates  

```{r, echo=F}
gender.ratio <-
  f.gen.dat.2014 %>%
  group_by(gender) %>%
  summarize(IUSSI_2014=sum(participants_2014)) %>%
  left_join((f.gen.dat.2018 %>%
              group_by(gender) %>%
              summarize(IUSSI_2018=sum(participants_2018))), by="gender")

library(kableExtra)
gender.ratio %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

> The, **estimated *female:male ratio* **among selected participants at **IUSSI 2014 was `r round(gender.ratio[1,2]/gender.ratio[2,2],2)`** and at **IUSSI 2018 was `r round(gender.ratio[1,3]/gender.ratio[2,3],2)`.**

***

## Participation by Career-stage

```{r, include=F}

## Exploring career stage of participants

titles.2014 <- dat.2014 %>% pull(title) %>% as.character()
titles.2018 <- dat.2018 %>% pull(degree) %>% as.character()

titles.2014 %>% unique()
titles.2018 %>% unique()

## Let's summarize participants by title (career-stage) 
  dat.2014 %>% 
  group_by(title) %>% 
  summarize(freq=n())
  
## Looks like it needs some clean-up
  # PIs - c("A/Prof","A/Professor","Ass/Prof","Associate Profes","Prof","Professor")
  # Graduate/Undergrads - c("Miss","Mr","Mrs","Ms")
  # Dr - most likely postdocs

career.2014 <- 
  dat.2014 %>% 
  mutate(degree = ifelse(title %in% c("A/Prof","A/Professor","Ass/Prof","Associate Profes","Prof","Professor"),
                         "Prof.",title)) %>% 
  mutate(degree = ifelse(title %in% c("Miss","Mr","Mrs","Ms"),
                         "Mr/Ms.",degree)) %>% 
  mutate(degree = ifelse(title %in% c("Dr"),
                         "Dr.",degree)) %>% 
  select(-title) %>% 
  select(degree, everything())

career.2014 <-
  career.2014 %>% 
  group_by(degree) %>% 
  summarise(freq = n()) %>% 
  filter(!degree %in% c("")) %>% # removed two with no assignment
  mutate(percent = round(freq/sum(freq)*100,1)) %>% 
  arrange(match(degree, c("Prof.","Dr.","Mr/Ms.")))


career.2018 <- 
  dat.2018 %>% 
  group_by(degree) %>% 
  summarize(freq=n()) %>% 
  filter(!degree %in% c("")) %>% # removed one with no assignment
  mutate(percent = round(freq/sum(freq)*100,1)) %>% 
  arrange(match(degree, c("Prof.","Dr.","Post-Doc","Graduate","Undergraduate","Other")))

```
   
* The data doesn't permit a comprehensive count of participants by their career stages since we have terms such as "Dr" that could indicate either a post-doc or a PI. 

* Additionally, the data for IUSSI 2014 has titles ("A/Prof","A/Professor","Ass/Prof","Associate Profes","Prof","Professor") that have been clubbed together as "Prof." and (Miss","Mr","Mrs","Ms") that are represented as "Mr/Ms."


```{r, echo=F}
career.2014 %>% 
  kable(caption = "IUSSI 2014") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

```{r, echo=F}
career.2018 %>% 
  kable(caption = "IUSSI 2018") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


> The data does show a **fairly good representation of graduate students** at both IUSSIs. 

However, in future, including the upcoming IUSSI 2022, it would be useful to come up with a standardized way to collect demography data during the registration process.

___

All code used for this document is available on my [Github](https://github.com/biplabendu/homepage/blob/master/IUSSI.Rmd).

___
